---
title: "Hands-on Ex 01"
author: "HU Yue"
---

## 1.3 Introducing ggplot

```{r}
library(readr)
exam_data <- read_csv("Exam_data.csv")

hist(exam_data$MATHS)
```

## Histogram of Maths scores

This code chunk creates a histogram for the variable `MATHS` in `exam_data`. It helps us understand the distribution of students' maths scores. The argument `bins = 10` groups the scores into 10 intervals. `color = "black"` sets the border color of each bar, and `fill = "grey"` sets the bar color. The title makes the chart easier to interpret.

```{r}
library(ggplot2)
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(
    bins = 10,
    boundary = 100,
    color = "black",
    fill = "grey"
  ) +
  ggtitle("Distribution of Maths scores")
```

## Why ggplot is prefered?

ggplot2 is more flexible, consistent, and easier to extend than base R graphics. For simple plots, base R is often quicker, but as the visualization becomes more complex, ggplot2 makes it much easier to control aesthetics, add layers, apply themes, and produce cleaner, more professional-looking charts. It also follows a clear grammar of graphics structure, so the code is usually more systematic and easier to read, modify, and reuse.

## 1.5 Essential Grammatical Elements in ggplot2: data

```{r}
ggplot(data = exam_data)
```

## 1.6 Essential Grammatical Elements in ggplot2: Aesthetic mappings

Code chunk below adds the aesthetic mapping into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS))
```

**Explanation**

This code chunk introduces the aesthetic mapping in `ggplot2`. The dataset used is `exam_data`, and the variable `MATHS` is mapped to the x-axis by using `aes(x = MATHS)`. As a result, the plot now shows the x-axis and its label. However, no geometric layer is added yet, so the chart area is still empty.

## 1.7.1 Geometric Objects: geom_bar

Code chunk below adds a geometric object into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE`. The dataset `exam_data` is first passed into `ggplot()`, and `aes(x = RACE)` maps the categorical variable `RACE` to the x-axis. The function `geom_bar()` then counts the number of observations in each category and displays them as bars. This makes it useful for showing the frequency distribution of a categorical variable.

## 1.7.2 Geometric Objects: geom_dotplot

Code chunk below adds a dot plot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

**Explanation**

This code chunk creates a dot plot for the variable `MATHS`. The dataset `exam_data` is passed into `ggplot()`, and `aes(x = MATHS)` maps the variable `MATHS` to the x-axis. The function `geom_dotplot()` represents each observation as a dot, making it useful for showing the distribution of a continuous variable. The argument `dotsize = 0.5` reduces the size of the dots so that the plot looks clearer and less crowded.

Code chunk below adds a dot plot and adjusts the scale of the y-axis.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth = 2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)
```

**Explanation**

This code chunk creates a dot plot for the variable `MATHS` and further refines its appearance. The function `geom_dotplot()` displays each observation as a dot. The argument `binwidth = 2.5` controls how the values are grouped along the x-axis, while `dotsize = 0.5` makes the dots smaller and easier to read. The function `scale_y_continuous(NULL, breaks = NULL)` removes the y-axis title and tick marks, which makes the plot cleaner because the y-axis in a dot plot is usually not the main focus.

## 1.7.3 Geometric Objects: geom_histogram()

Code chunk below adds a histogram into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram()
```

**Explanation**

This code chunk creates a histogram for the variable `MATHS`. The dataset `exam_data` is first passed into `ggplot()`, and `aes(x = MATHS)` maps the variable `MATHS` to the x-axis. The function `geom_histogram()` then groups the values of `MATHS` into intervals and displays their frequencies as bars. This makes it useful for showing the distribution of a continuous variable.

## 1.7.4 Modifying a geometric object by changing geom()

Code chunk below customizes the histogram by changing the number of bins and the bar appearance.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20,            
                 color = "black",      
                 fill = "lightblue")
```

**Explanation**

This code chunk creates a histogram for the variable `MATHS` and customizes its appearance. The function `geom_histogram()` groups the data into 20 bins, so the distribution can be shown in more detail. The argument `color = "black"` adds black borders to each bar, while `fill = "lightblue"` sets the bar color to light blue. These settings make the chart clearer and easier to interpret.

## 1.7.5 Modifying a geometric object by changing aes()

Code chunk below adds the `fill` aesthetic to the histogram.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           fill = GENDER)) +
  geom_histogram(bins = 20, 
                 color = "grey30")
```

**Explanation**

This code chunk creates a histogram of the variable `MATHS` and uses `GENDER` as the fill aesthetic. The variable `MATHS` is mapped to the x-axis, while `GENDER` is mapped to the fill color of the bars. As a result, the histogram shows how the distribution of maths scores differs by gender. The argument `bins = 20` divides the data into 20 intervals, and `color = "grey30"` adds dark grey borders to the bars for clearer separation.

## 1.7.6 Geometric Objects: geom-density()

Code chunk below adds a density curve into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_density()
```

**Explanation**

This code chunk creates a density plot for the variable `MATHS`. The dataset `exam_data` is passed into `ggplot()`, and `aes(x = MATHS)` maps the variable `MATHS` to the x-axis. The function `geom_density()` then draws a smoothed density curve to show the overall distribution of the maths scores. Compared with a histogram, a density plot provides a smoother view of the shape of the distribution.

Code chunk below adds the `colour` aesthetic to the density plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

**Explanation**

This code chunk creates a density plot for the variable `MATHS` and uses `GENDER` as the colour aesthetic. The variable `MATHS` is mapped to the x-axis, while `GENDER` is mapped to the line colour. As a result, separate density curves are drawn for each gender, allowing us to compare the distribution of maths scores between groups more clearly.

## 1.7.7 Geometric Objects: geom_boxplot

Code chunk below adds a boxplot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS,       
           x = GENDER)) +    
  geom_boxplot()
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER`. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_boxplot()` summarizes the distribution of maths scores for each gender by showing the median, quartiles, and potential outliers. This makes it useful for comparing the center and spread of scores across groups.

Code chunk below adds a boxplot with notches into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, 
           x = GENDER)) +
  geom_boxplot(notch = TRUE)
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER` and adds notches to the boxes. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_boxplot()` summarizes the distribution of maths scores for each gender by showing the median, quartiles, and possible outliers. The argument `notch = TRUE` adds a notch around the median, which provides a visual guide for comparing the medians between groups.

## 1.7.8 Geometric Objects: geom_violin

Code chunk below adds a violin plot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, 
           x = GENDER)) +
  geom_violin()
```

**Explanation**

This code chunk creates a violin plot of `MATHS` grouped by `GENDER`. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_violin()` shows the distribution and density of maths scores for each gender. Compared with a boxplot, a violin plot provides more information about the shape of the distribution.

## 1.7.9 Geometric Objects: geom_point()

Code chunk below adds a scatter plot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point()
```

**Explanation**

This code chunk creates a scatter plot using `MATHS` and `ENGLISH`. The variable `MATHS` is mapped to the x-axis, while `ENGLISH` is mapped to the y-axis. The function `geom_point()` plots each observation as a point. This type of chart is useful for examining the relationship between two numerical variables and identifying patterns, trends, or possible outliers.

## 1.7.10 geom objects can be combined

Code chunk below combines a boxplot with jittered points.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, 
           x = GENDER)) +
  geom_boxplot() +                    
  geom_point(position = "jitter", 
             size = 0.5)
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER` and then adds individual data points on top of it. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. `geom_boxplot()` summarizes the distribution by showing the median, quartiles, and possible outliers. `geom_point(position = "jitter", size = 0.5)` adds the raw observations and slightly spreads them horizontally so that overlapping points can be seen more clearly. This makes the plot more informative because it shows both the summary statistics and the actual data values.

## 1.8 Essential Grammatical Elements in ggplot2: stat

## 1.8.1 Working with stat()

Code chunk below adds a boxplot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, x = GENDER)) +
  geom_boxplot()
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER`. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_boxplot()` summarizes the distribution of maths scores for each gender by showing the median, quartiles, and possible outliers. This plot is useful for comparing the center and spread of scores across groups.

## 1.8.2 Working with stat - the stat_summary() method

Code chunk below adds a boxplot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, x = GENDER)) +
  geom_boxplot()
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER`. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_boxplot()` summarizes the distribution of maths scores for each gender by showing the median, quartiles, and possible outliers. This plot is useful for comparing the center and spread of scores across groups.

## 1.8.3 Working with stat - the geom() method

Code chunk below adds a boxplot into the plot.

```{r}
ggplot(data = exam_data, 
       aes(y = MATHS, x = GENDER)) +
  geom_boxplot()
```

**Explanation**

This code chunk creates a boxplot of `MATHS` grouped by `GENDER`. The variable `GENDER` is mapped to the x-axis, while `MATHS` is mapped to the y-axis. The function `geom_boxplot()` summarizes the distribution of maths scores for each gender by showing the median, quartiles, and possible outliers. This plot is useful for comparing the center and spread of scores across groups.

## 1.8.4 Adding a best fit curve on a scatterplot

Code chunk below creates a scatter plot and adds a smoothing line.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(size = 0.5)
```

**Explanation**

geom_point() shows the individual observations, while geom_smooth() adds a smooth trend line. This helps us see the overall relationship between MATHS and ENGLISH more clearly.

Code chunk below creates a scatter plot and adds a linear regression line.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              linewidth = 0.5)
```

**Explanation**

method = lm tells geom_smooth() to use a linear model, so the fitted line is a straight regression line rather than a general smooth curve.

## 1.9.1 Working with facet_wrap()

Code chunk below creates a histogram and splits it into separate panels by `CLASS`.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20) +
  facet_wrap(~ CLASS)
```

**Explanation**

This code chunk creates a histogram of the variable `MATHS` and then uses `facet_wrap(~ CLASS)` to split the plot into separate panels according to `CLASS`. The variable `MATHS` is mapped to the x-axis, and `geom_histogram(bins = 20)` groups the scores into 20 bins to show their distribution. `facet_wrap(~ CLASS)` produces a separate histogram for each class, making it easier to compare the distribution of maths scores across different classes.

## 1.9.2 facet_grid() function

Code chunk below creates a histogram and splits it into separate panels by `CLASS`.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

**Explanation**

This code chunk creates a histogram of the variable `MATHS` and then uses `facet_grid(~ CLASS)` to divide the plot into separate panels according to `CLASS`. The variable `MATHS` is mapped to the x-axis, and `geom_histogram(bins = 20)` groups the maths scores into 20 bins to show their distribution. `facet_grid(~ CLASS)` produces one panel for each class, which makes it easier to compare the distribution of maths scores across different classes.

## 1.10.1 Working with Coordinate

Code chunk below adds a bar chart into the plot.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE`. The dataset `exam_data` is first passed into `ggplot()`, and `aes(x = RACE)` maps the categorical variable `RACE` to the x-axis. The function `geom_bar()` then counts the number of observations in each category and displays them as bars. This makes it useful for showing the frequency distribution of a categorical variable.

Code chunk below creates a bar chart and flips the coordinates.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar() +
  coord_flip()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE` and then flips the axes by using `coord_flip()`. The variable `RACE` is mapped to the x-axis, and `geom_bar()` counts the number of observations in each category and displays them as bars. `coord_flip()` switches the x-axis and y-axis, so the bars become horizontal instead of vertical. This is useful when category labels are long or when a horizontal bar chart is easier to read.

## 1.10.2 Changing the y- and x-axis range

Code chunk below creates a scatter plot and adds a linear regression line.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5)
```

**Explanation**

This code chunk creates a scatter plot using `MATHS` and `ENGLISH`, and then adds a fitted linear regression line. The variable `MATHS` is mapped to the x-axis, while `ENGLISH` is mapped to the y-axis. `geom_point()` plots each observation as a point, allowing us to see the relationship between the two variables. `geom_smooth(method = lm, size = 0.5)` fits a straight line based on a linear model and overlays it on the scatter plot. This helps us assess whether there is a linear relationship between maths and English scores.

Code chunk below creates a scatter plot with a linear regression line and limits the visible plotting range.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              size = 0.5) +  
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100))
```

**Explanation**

This code chunk creates a scatter plot using `MATHS` and `ENGLISH`, and then adds a fitted linear regression line. The variable `MATHS` is mapped to the x-axis, while `ENGLISH` is mapped to the y-axis. `geom_point()` plots each observation as a point, allowing us to examine the relationship between the two variables. `geom_smooth(method = lm, size = 0.5)` adds a straight regression line based on a linear model. `coord_cartesian(xlim = c(0, 100), ylim = c(0, 100))` restricts the visible range of both axes to values between 0 and 100, making the plot focus on the score range more clearly.

## 1.11.1 Working with theme

Code chunk below creates a horizontal bar chart and applies the gray theme.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE`, flips the coordinates, and applies a theme. The variable `RACE` is mapped to the x-axis, and `geom_bar()` counts the number of observations in each category and displays them as bars. `coord_flip()` switches the axes so that the bars become horizontal, which can improve readability. `theme_gray()` applies the default gray theme in `ggplot2`, which controls the overall appearance of the plot, including the background, grid lines, and text style.

Code chunk below creates a horizontal bar chart and applies the classic theme.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE`, flips the coordinates, and applies a theme. The variable `RACE` is mapped to the x-axis, and `geom_bar()` counts the number of observations in each category and displays them as bars. `coord_flip()` switches the axes so that the bars become horizontal, which makes the chart easier to read. `theme_classic()` changes the overall appearance of the plot by removing the gray background and grid lines, giving the chart a cleaner and simpler style.

Code chunk below creates a horizontal bar chart and applies the minimal theme.

```{r}
ggplot(data = exam_data, 
       aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```

**Explanation**

This code chunk creates a bar chart for the variable `RACE`, flips the coordinates, and applies a theme. The variable `RACE` is mapped to the x-axis, and `geom_bar()` counts the number of observations in each category and displays them as bars. `coord_flip()` switches the axes so that the bars become horizontal, which improves readability. `theme_minimal()` changes the overall appearance of the plot by using a clean and simple style with fewer visual distractions, making the chart look more modern and uncluttered.
