---
title: "Take-Home_Ex02"
author: "HU YUE"
date: "2026-05-27"
date-modified: "last-modified"
format:
  html:
    theme: cosmo
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: true
    code-tools: true
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

# 1. Introduction

TenantThread is a proptech company whose internal communications became the focus of a crisis investigation after merger-related information entered the public domain before the expected communication process was fully controlled. The case involves a communications embargo around the CivicLoom–TenantThread merger, internally referred to as Project HarborCrest, alongside broader pressure from media scrutiny, regulatory concerns, social media activity, and client-facing risk. The dataset describes a multi-agent corporate communications environment in which seven AI agents interacted across multiple channels and made real-time decisions about what should or should not be published. :contentReference[oaicite:0]{index="0"}

The purpose of this project is to reconstruct TenantThread’s internal communications during the two weeks leading up to the breach. Rather than treating the breach as a single isolated event, this investigation examines how information, pressure, and decision-making evolved over time across internal messages, agent deliberations, and public posts. The central question is:

> Did TenantThread’s team deliberately leak the merger, or did the AI-managed communication system break down under pressure?

To answer this question, the project develops a single-page visual analytics website. The analysis will first examine the overall structure of the data, then reconstruct the communication timeline, identify key actors and communication pathways, assess signs of system pressure, trace merger-related information, and compare internal communications with public-facing posts. The final assessment will be based on visual evidence rather than assumptions made before the data is analysed.

This website is organised as a visual investigation. Each section contributes to one part of the evidence chain: what happened, who was involved, how information moved, whether the system showed signs of overload, and whether public posts reflected prior internal communications. Together, these components support a structured judgement on whether the breach is more consistent with deliberate disclosure or system breakdown.

# Executive Summary

This visual analytics report evaluates two competing explanations:

| Hypothesis | Explanation |
|------------------------------------|------------------------------------|
| Deliberate Leak | Someone intentionally moved confidential merger information into public-facing communication. |
| System Breakdown | The AI-managed communication system failed under pressure, causing confidential information to cross into public channels. |

The analysis proceeds through five stages:

1.  Reconstruct the two-week event timeline.
2.  Identify key actors and communication pathways.
3.  Detect abnormal pressure and breakdown signals.
4.  Trace merger-related information across internal and public channels.
5.  Compare internal deliberations with public posts.

# 2. Data Overview

This section provides an overview of the dataset used in this investigation. The purpose is to understand the structure, coverage, and analytical potential of the data before moving into detailed visual analysis.

The dataset contains TenantThread’s multi-agent corporate communication records during the two weeks leading up to the information release. It includes internal communications, agent-level decision processes, public-facing posts, environmental context, market signals, media events, and social media conditions.

The original JSON file is organised around communication rounds. Each round contains a timestamp, environmental context, communication records, and participating agents. This structure allows the investigation to connect what agents said, what they were reacting to, and what external conditions existed at each point in time.

```{r}
library(tidyverse)
library(jsonlite)
library(lubridate)
library(stringr)
library(DT)
```

## Read the JSON file

```{r}
mc1_raw <- fromJSON(
  "data/MC1_final_00.json",
  flatten = TRUE
)

rounds <- mc1_raw$rounds
```

## Create communication-level dataset

```{r}
communications <- rounds %>%
  select(
    round_hour = hour,
    communications
  ) %>%
  unnest(communications) %>%
  mutate(
    timestamp = ymd_hms(timestamp),
    round_hour = ymd_hms(round_hour),
    date = as_date(timestamp),
    hour_of_day = hour(timestamp),
    content_lower = str_to_lower(content),
    is_public = message_type == "public_post"
  )
```

## Data source summary

```{r}
data_summary <- tibble(
  Dataset = c("Communication records", "Communication rounds"),
  Description = c(
    "Individual messages, posts, and agent communications",
    "Hourly crisis context and grouped communication events"
  ),
  Records = c(
    nrow(communications),
    nrow(rounds)
  )
)

datatable(
  data_summary,
  rownames = FALSE,
  options = list(pageLength = 5)
)
```

## Time Coverage

```{r}
time_coverage <- communications %>%
  summarise(
    start_time = min(timestamp, na.rm = TRUE),
    end_time = max(timestamp, na.rm = TRUE),
    number_of_days = as.numeric(max(date, na.rm = TRUE) - min(date, na.rm = TRUE)) + 1,
    total_messages = n()
  )

datatable(
  time_coverage,
  rownames = FALSE,
  options = list(dom = "t")
)
```

## Agents Overview

```{r}
agent_overview <- communications %>%
  count(agent_id, agent_role, agent_label, sort = TRUE) %>%
  rename(
    Agent_ID = agent_id,
    Role = agent_role,
    Label = agent_label,
    Message_Count = n
  )

datatable(
  agent_overview,
  rownames = FALSE,
  options = list(pageLength = 10)
)
```

## Communication Channels Overview

```{r}
channel_overview <- communications %>%
  count(channel, message_type, sort = TRUE) %>%
  rename(
    Channel = channel,
    Message_Type = message_type,
    Count = n
  )

datatable(
  channel_overview,
  rownames = FALSE,
  options = list(pageLength = 10)
)
```

## Simple Overview

```{r}
communications %>%
  count(message_type) %>%
  ggplot(aes(x = reorder(message_type, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Communications by Message Type",
    x = "Message Type",
    y = "Number of Records"
  ) +
  theme_minimal()
```

Overall, this overview establishes the basic structure of the dataset: when the communications occurred, which agents participated, and which channels were used. The next section moves from data description to data preparation, where the raw nested JSON structure is transformed into analysis-ready tables for timeline, network, pressure, and leak-tracing visualisations.

# 3. Data Preparation and Integration

This section prepares the raw JSON-derived data for later visual analysis. The data description file is used only as a reference for understanding the dataset structure, while the actual data preparation is based on the fields available in the JSON file.

## 3.1 Schema Validation

Before transforming the data, the actual structure of the JSON object is inspected. This step ensures that the following extraction process is based on the available fields in the dataset.

```{r}
names(rounds)
```

```{r}
schema_check <- tibble(
  field = c(
    "hour",
    "environment_context",
    "communications",
    "participants"
  ),
  present_in_actual_json = c(
    "hour" %in% names(rounds),
    "environment_context" %in% names(rounds),
    "communications" %in% names(rounds),
    "participants" %in% names(rounds)
  )
)

DT::datatable(
  schema_check,
  rownames = FALSE,
  options = list(dom = "t")
)
```

## 3.2 Extract Communication Records

```{r}
communications <- rounds %>%
  select(
    round_hour = hour,
    communications
  ) %>%
  unnest(communications)

glimpse(communications)
```

## 3.3 Standardise Communication Records

```{r}
communications_clean <- communications %>%
  mutate(
    timestamp = ymd_hms(timestamp),
    round_hour = ymd_hms(round_hour),
    date = as_date(timestamp),
    hour_of_day = hour(timestamp),
    day_of_week = wday(timestamp, label = TRUE),
    content_lower = str_to_lower(content),
    content_length = str_count(content, "\\S+"),
    is_public = message_type == "public_post"
  )

glimpse(communications_clean)
```

## 3.4 Construct Analytical Variables

```{r}
communications_clean <- communications_clean %>%
  mutate(
    contains_merger = str_detect(
      content_lower,
      "merger|acquisition|civicloom|harborcrest|project harborcrest|deal|transaction"
    ),
    
    contains_confidential = str_detect(
      content_lower,
      "confidential|embargo|private|internal only|not public|restricted"
    ),
    
    contains_urgency = str_detect(
      content_lower,
      "urgent|asap|immediately|deadline|pressure|crisis|risk|escalate|failed|error"
    ),
    
    communication_scope = case_when(
      is_public ~ "Public",
      channel == "one_on_one_chat" ~ "Private / One-on-One",
      TRUE ~ "Internal"
    )
  )
```

## 3.5 Extract Environmental Context

```{r}
environment_context <- rounds %>%
  transmute(
    round_hour = ymd_hms(hour),
    event_narrative = environment_context.event_narrative,
    event_headline = environment_context.event_headline,
    stock_price = environment_context.market_snapshot.stock_price,
    percent_change = environment_context.market_snapshot.percent_change,
    market_sentiment = environment_context.market_snapshot.sentiment,
    social_state = environment_context.social_state
  )

glimpse(environment_context)
```

## 3.6 Extract Participant Data

```{r}
participants <- rounds %>%
  select(
    round_hour = hour,
    participants
  ) %>%
  unnest(participants) %>%
  mutate(
    round_hour = ymd_hms(round_hour)
  )

glimpse(participants)
```

## 3.7 Integrate Communication and Context Data

```{r}
communications_integrated <- communications_clean %>%
  left_join(
    environment_context,
    by = "round_hour"
  )

glimpse(communications_integrated)
```

## 3.8 Export Processed Tables

```{r}
dir.create("data/processed", recursive = TRUE, showWarnings = FALSE)

write_csv(
  communications_clean,
  "data/processed/communications_clean.csv"
)

write_csv(
  environment_context,
  "data/processed/environment_context.csv"
)

write_csv(
  participants,
  "data/processed/participants.csv"
)

write_csv(
  communications_integrated,
  "data/processed/communications_integrated.csv"
)
```

## 3.9 Data Preparation Summary

```{r}
prep_summary <- tibble(
  prepared_table = c(
    "communications_clean",
    "environment_context",
    "participants",
    "communications_integrated"
  ),
  unit_of_analysis = c(
    "One row per communication record",
    "One row per crisis round",
    "One row per agent per round",
    "One row per communication record with contextual variables"
  ),
  main_use = c(
    "Timeline reconstruction, keyword analysis, public post analysis",
    "Crisis context and event background",
    "Agent participation overview",
    "Integrated analysis combining messages with crisis context"
  ),
  records = c(
    nrow(communications_clean),
    nrow(environment_context),
    nrow(participants),
    nrow(communications_integrated)
  )
)

DT::datatable(
  prep_summary,
  rownames = FALSE,
  options = list(pageLength = 5)
)
```

After preparation, the nested JSON structure has been transformed into analysis-ready tables. These processed datasets will support the subsequent timeline reconstruction, communication flow analysis, system pressure analysis, merger information tracing, and public post comparison.

# 4. Timeline Reconstruction

This section reconstructs the communication sequence during the two weeks leading up to the breach. The purpose is to understand when different types of communications occurred, how activity evolved over time, and where public posts appeared within the broader internal communication flow.

## 4.1 Prepare Timeline Data

```{r}
timeline_data <- communications_integrated %>%
  mutate(
    source_group = case_when(
      is_public ~ "Public Post",
      channel == "one_on_one_chat" ~ "One-on-One Chat",
      channel == "comms_huddle" ~ "Comms Huddle",
      TRUE ~ "Other Internal Communication"
    ),
    short_content = str_trunc(content, width = 120)
  ) %>%
  arrange(timestamp)
```

```{r}
breach_time <- ymd_hms("2046-06-05 17:00:00")
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")

timeline_data <- timeline_data %>%
  mutate(
    breach_window = case_when(
      timestamp < breach_time ~ "Before 5:00 PM Breach",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5:00–6:00 PM Embargo Breach Window",
      timestamp >= embargo_lift_time ~ "After 6:00 PM Embargo Lift"
    )
  )
```

## 4.2 Overall Communication Timeline

```{r}
ggplot(
  timeline_data,
  aes(
    x = timestamp,
    y = source_group,
    color = source_group
  )
) +
  geom_point(alpha = 0.75, size = 2) +
  labs(
    title = "Communication Timeline Before the Breach",
    subtitle = "Each point represents one communication record",
    x = "Timestamp",
    y = "Communication Type",
    color = "Communication Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

The timeline indicates that TenantThread’s communication activity continued throughout the two-week period, with Comms Huddle, One-on-One Chat, and Public Posts appearing repeatedly. This suggests that public-facing communication occurred alongside ongoing internal discussion rather than as a single isolated event. “Other Internal Communication” is much less frequent and appears only on selected dates. The clustering of records near the end of the period suggests that later sections should examine whether communication density, merger-related content, or urgency signals increased closer to the breach.

## 4.3 Daily Communication Volume

```{r}
daily_volume <- timeline_data %>%
  count(date, source_group)

ggplot(
  daily_volume,
  aes(
    x = date,
    y = n,
    fill = source_group
  )
) +
  geom_col() +
  labs(
    title = "Daily Communication Volume",
    subtitle = "Number of communication records by day and communication type",
    x = "Date",
    y = "Number of Communications",
    fill = "Communication Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

Daily communication volume remains relatively stable during most of the two-week period, but rises sharply in early June. The final spike reaches almost 500 communication records, far above the earlier daily levels. Importantly, this increase includes multiple communication types rather than only public posts, suggesting a broader escalation of internal and external communication activity. This pattern provides an initial indication that the breach occurred during a high-pressure communication period, which will be examined further in the system pressure analysis.

## 4.4 Merger-Related Communication Timeline

```{r}
merger_spread_heatmap <- timeline_data %>%
  filter(contains_merger == TRUE) %>%
  count(date, agent_label, name = "merger_count")

ggplot(
  merger_spread_heatmap,
  aes(
    x = date,
    y = agent_label,
    fill = merger_count
  )
) +
  geom_tile(
    color = "white",
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = ifelse(merger_count > 0, merger_count, "")),
    size = 3
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkblue"
  ) +
  labs(
    title = "Spread of Merger-Related Communication Across Agents",
    subtitle = "Darker cells indicate more merger-related communications by agent and date",
    x = "Date",
    y = "Agent",
    fill = "Merger Count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

Merger-related communications are sparse in the earlier period but become much more concentrated near early June. The later cluster involves more agents, including Legal, PR, Social Manager, PR Intern, Judge, and Intern agents. This suggests that merger-related information became more widely circulated close to the breach period, which makes the later information-flow and public-post comparison especially important.

## 4.5 Key Signal Trends Over Time

```{r}
signal_escalation_heatmap <- timeline_data %>%
  group_by(date) %>%
  summarise(
    `Merger-Related` = sum(contains_merger, na.rm = TRUE),
    `Confidentiality-Related` = sum(contains_confidential, na.rm = TRUE),
    `Urgency-Related` = sum(contains_urgency, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(
      `Merger-Related`,
      `Confidentiality-Related`,
      `Urgency-Related`
    ),
    names_to = "signal_type",
    values_to = "message_count"
  )

ggplot(
  signal_escalation_heatmap,
  aes(
    x = date,
    y = signal_type,
    fill = message_count
  )
) +
  geom_tile(
    color = "white",
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = ifelse(message_count > 0, message_count, "")),
    size = 3
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkred"
  ) +
  labs(
    title = "Key Signal Escalation Over Time",
    subtitle = "Merger, confidentiality, and urgency signals by date",
    x = "Date",
    y = "Signal Type",
    fill = "Message Count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

The graph shows that merger, confidentiality, and urgency signals stay low for most of the period, then surge sharply in early June. Merger-related signals increase the most, followed by urgency and confidentiality signals. This suggests that the breach period coincided with both wider merger discussion and heightened communication pressure.

## 4.6 Breach Window Signal Intensity on June 5

This visual zooms into June 5, 2046, the day of the inappropriate release. The vertical reference lines mark the approximate breach time at 5:00 PM and the formal embargo lift at 6:00 PM. This allows the analysis to separate behavior before the breach, during the embargo breach window, and after the embargo was formally lifted.

```{r}
breach_time <- ymd_hms("2046-06-05 17:00:00")
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")

breach_day_data <- timeline_data %>%
  filter(date == as_date("2046-06-05")) %>%
  mutate(
    breach_phase = case_when(
      timestamp < breach_time ~ "Before 5:00 PM",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5:00–6:00 PM Breach Window",
      timestamp >= embargo_lift_time ~ "After 6:00 PM Embargo Lift"
    ),
    breach_phase = factor(
      breach_phase,
      levels = c(
        "Before 5:00 PM",
        "5:00–6:00 PM Breach Window",
        "After 6:00 PM Embargo Lift"
      )
    )
  )

day_start <- min(breach_day_data$timestamp, na.rm = TRUE)
day_end <- max(breach_day_data$timestamp, na.rm = TRUE)

phase_duration <- tibble(
  breach_phase = factor(
    c(
      "Before 5:00 PM",
      "5:00–6:00 PM Breach Window",
      "After 6:00 PM Embargo Lift"
    ),
    levels = c(
      "Before 5:00 PM",
      "5:00–6:00 PM Breach Window",
      "After 6:00 PM Embargo Lift"
    )
  ),
  duration_hours = c(
    as.numeric(difftime(breach_time, day_start, units = "hours")),
    1,
    as.numeric(difftime(day_end, embargo_lift_time, units = "hours"))
  )
)

breach_phase_summary <- breach_day_data %>%
  group_by(breach_phase) %>%
  summarise(
    `Public + Merger` = sum(is_public & contains_merger, na.rm = TRUE),
    `Public Posts` = sum(is_public, na.rm = TRUE),
    `Merger-Related` = sum(contains_merger, na.rm = TRUE),
    `Confidentiality-Related` = sum(contains_confidential, na.rm = TRUE),
    `Urgency-Related` = sum(contains_urgency, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    phase_duration,
    by = "breach_phase"
  ) %>%
  pivot_longer(
    cols = c(
      `Public + Merger`,
      `Public Posts`,
      `Merger-Related`,
      `Confidentiality-Related`,
      `Urgency-Related`
    ),
    names_to = "signal_type",
    values_to = "message_count"
  ) %>%
  mutate(
    intensity_per_hour = message_count / duration_hours,
    signal_type = factor(
      signal_type,
      levels = c(
        "Public + Merger",
        "Public Posts",
        "Merger-Related",
        "Confidentiality-Related",
        "Urgency-Related"
      )
    )
  )

ggplot(
  breach_phase_summary,
  aes(
    x = intensity_per_hour,
    y = signal_type,
    fill = breach_phase
  )
) +
  geom_col(
    position = position_dodge(width = 0.75),
    width = 0.65
  ) +
  labs(
    title = "Breach Window Signal Intensity on June 5",
    subtitle = "Signal counts standardised as messages per hour across key embargo phases",
    x = "Messages per Hour",
    y = "Signal Type",
    fill = "Breach Phase"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

The chart shows that merger-related and public-merger signals were already present before the 6:00 PM embargo lift. The 5:00–6:00 PM breach window has a clear increase in merger-related and public-facing merger activity, which suggests that embargoed information had started moving into public communication before the authorised release time. After 6:00 PM, all signal types rise further, but the key issue is that public merger-related communication was not zero before the embargo expired.

## 4.7 Embargo Window Communication Breakdown

The merger announcement was under embargo until 6:00 PM on June 5. Therefore, this section focuses specifically on communication before the embargo lift, especially merger-related messages and public-facing posts before 6:00 PM.

```{r}
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

pre_embargo_data <- communications_integrated %>%
  filter(
    date == as_date("2046-06-05"),
    timestamp < embargo_lift_time
  ) %>%
  mutate(
    embargo_phase = case_when(
      timestamp < breach_time ~ "Before 5:00 PM",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5:00–6:00 PM Breach Window"
    ),
    embargo_risk_type = case_when(
      is_public & contains_merger ~ "Public + Merger",
      is_public ~ "Public Only",
      contains_merger ~ "Internal/Private Merger",
      contains_confidential ~ "Confidentiality",
      contains_urgency ~ "Urgency",
      TRUE ~ "Other"
    )
  )
```

```{r}
pre_embargo_merger_by_agent <- pre_embargo_data %>%
  filter(contains_merger == TRUE) %>%
  count(agent_label, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(total = sum(message_count)) %>%
  ungroup()

ggplot(
  pre_embargo_merger_by_agent,
  aes(
    x = reorder(agent_label, total),
    y = message_count,
    fill = communication_scope
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Merger-Related Communications Before Embargo Lift",
    subtitle = "June 5 before 6:00 PM, grouped by agent and communication scope",
    x = "Agent",
    y = "Number of Merger-Related Communications",
    fill = "Communication Scope"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

```{r}
pre_embargo_merger_channel <- pre_embargo_data %>%
  filter(contains_merger == TRUE) %>%
  count(channel, message_type, name = "message_count") %>%
  arrange(desc(message_count))

ggplot(
  pre_embargo_merger_channel,
  aes(
    x = reorder(channel, message_count),
    y = message_count,
    fill = message_type
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Channels Carrying Merger-Related Information Before 6:00 PM",
    subtitle = "Embargoed merger information by channel and message type",
    x = "Channel",
    y = "Number of Merger-Related Communications",
    fill = "Message Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

```{r}
public_merger_before_embargo <- pre_embargo_data %>%
  filter(
    is_public == TRUE,
    contains_merger == TRUE
  ) %>%
  arrange(timestamp) %>%
  transmute(
    Time = timestamp,
    Phase = embargo_phase,
    Agent = agent_label,
    Channel = channel,
    Message_Type = message_type,
    Content = str_trunc(content, width = 220)
  )

DT::datatable(
  public_merger_before_embargo,
  rownames = FALSE,
  options = list(
    pageLength = 8,
    lengthMenu = c(8, 15, 30),
    scrollX = TRUE,
    scrollY = "320px",
    autoWidth = FALSE,
    columnDefs = list(
      list(width = "130px", targets = 0),
      list(width = "160px", targets = 1),
      list(width = "150px", targets = 2),
      list(width = "120px", targets = 3),
      list(width = "130px", targets = 4),
      list(width = "560px", targets = 5)
    )
  )
) %>%
  DT::formatStyle(
    "Content",
    `white-space` = "normal",
    `font-size` = "12px",
    `line-height` = "1.25"
  )
```

```{r}
embargo_risk_summary <- pre_embargo_data %>%
  summarise(
    total_pre_embargo_messages = n(),
    merger_related_before_6pm = sum(contains_merger, na.rm = TRUE),
    public_posts_before_6pm = sum(is_public, na.rm = TRUE),
    public_merger_posts_before_6pm = sum(is_public & contains_merger, na.rm = TRUE),
    one_on_one_merger_before_6pm = sum(channel == "one_on_one_chat" & contains_merger, na.rm = TRUE),
    comms_huddle_merger_before_6pm = sum(channel == "comms_huddle" & contains_merger, na.rm = TRUE)
  )

DT::datatable(
  embargo_risk_summary,
  rownames = FALSE,
  options = list(dom = "t", scrollX = TRUE)
)
```

The embargo-window breakdown shows that merger-related information was already active before the official 6:00 PM embargo lift. Legal-Agent and Social-Manager-Agent contributed the largest volume of merger-related communications before 6:00 PM, with activity spread across internal, private, and public scopes. This suggests that merger information was not confined to a narrow legal or executive channel.

The channel breakdown is especially important: merger-related information appeared not only in `comms_huddle` and `one_on_one_chat`, but also in `personal_post`, `official_post`, and `anonymous_post`. Because the merger announcement was still under embargo before 6:00 PM, the presence of merger-related content in public-facing channels provides direct evidence that the embargo boundary had already been crossed before the authorised release time.

## 4.8 Slack Screenshot as the Turning Point from Inference to Direct Leakage

```{r}
library(tidyverse)
library(lubridate)
library(ggrepel)

slack_timeline <- tribble(
  ~time, ~event, ~phase, ~severity,
  "2046-06-05 11:37:00",
  "CivicLoom speculation may overtake the ResidentIQ story before noon",
  "Market inference",
  2,
  "2046-06-05 11:41:00",
  "Legal: internal Slack screenshots are reaching Said-it within 20 minutes",
  "Internal leak",
  4,
  "2046-06-05 11:51:00",
  "Social-Manager: Employee Slack says 'the merger saves all our jobs'",
  "Direct public evidence",
  5,
  "2046-06-05 11:53:00",
  "Legal: Slack leak supports Section 8.1 confidentiality breach argument",
  "Legal breach",
  5,
  "2046-06-05 17:00:00",
  "SaltWind publishes TenantThread-CivicLoom merger story",
  "Public confirmation",
  5
) %>%
  mutate(time = ymd_hms(time))

ggplot(slack_timeline, aes(x = time, y = severity)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 3) +
  geom_label_repel(
    aes(label = event),
    size = 3,
    max.overlaps = Inf,
    label.size = 0.2
  ) +
  scale_y_continuous(
    breaks = c(2, 4, 5),
    labels = c("Inference", "Internal leak", "Confirmed breach")
  ) +
  labs(
    title = "Slack Screenshot Escalation Timeline",
    subtitle = "The Slack evidence marks the shift from market inference to direct internal leakage",
    x = "Time on 5 June 2046",
    y = "Disclosure stage"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 9)
  )
```

This timeline adds event-level evidence to the communication-volume timeline. The earlier part of the crisis was dominated by market inference, including speculation around CivicLoom and ResidentIQ. However, the Slack screenshot changed the nature of the breach. Legal reported that internal Slack screenshots were reaching Said-it within approximately 20 minutes, while Social-Manager later listed an employee Slack post saying “the merger saves all our jobs” as a confirmed public data point. This means the pre-6 PM exposure was not only a product of external speculation; it included direct internal leakage.

## 4.9 Key Findings from Timeline Reconstruction

- **The breach occurred during a clear escalation phase, not during routine communication.**\
  Communication activity remained moderate for most of the two-week period, but surged sharply on June 5. Internal, private, public, and other communication types all increased at the same time, indicating that the organisation had entered crisis communication mode rather than a normal coordination period.

- **Merger-related information became highly concentrated near the breach window.**\
  The agent-date heatmap shows that merger-related communication was sparse earlier in the period but became highly concentrated on June 5, especially around Legal-Agent and Social-Manager-Agent. This suggests that merger information had moved beyond a narrow legal-control context and was being actively discussed, monitored, and managed across multiple communication spaces.

- **Urgency, merger, and confidentiality signals intensified together.**\
  The signal escalation heatmap shows a strong concentration of urgency, merger-related, and confidentiality-related signals on June 5. This indicates that the release happened during a period when sensitive merger information, legal risk, and communication pressure were escalating simultaneously.

- **The Slack screenshot marks the turning point from market inference to direct internal leakage.**\
  Before the formal embargo lift, public discussion was not based only on external speculation or weak signals. The dataset records an employee Slack message stating that "the merger saves all our jobs" appearing on Said-it/r/PropTech, while Legal later described internal Slack screenshots reaching Said-it within about 20 minutes and treated this as a compromised information perimeter. This means the breach included direct internal leakage, not merely public inference.

- **The 5:00–6:00 PM window is critical because the embargo was still active.**\
  The breach-window comparison shows that public-merger signals were already present before the official 6:00 PM embargo lift. This is direct evidence that embargoed merger information had entered public-facing communication before it was formally authorised.

- **Pre-embargo merger communication was not limited to private or legal channels.**\
  Before 6:00 PM, merger-related information appeared across internal, private, and public scopes, and across channels including `side_huddle`, `one_on_one_chat`, `comms_huddle`, `personal_post`, `official_post`, and `anonymous_post`. This broad channel distribution shows that the confidentiality boundary had weakened before the official release.

- **Overall, the timeline evidence supports a mixed leakage pathway rather than a simple system-breakdown explanation.**\
  The inappropriate release appears to have developed in stages. First, market participants and external observers began reconstructing the merger from partial signals. Second, the Slack screenshot introduced direct internal evidence into the public domain. Third, Legal and communication agents shifted toward accelerated disclosure and narrative control. Therefore, the timeline does not support a clean binary conclusion of either deliberate leak or accidental system failure. The stronger interpretation is that an internal confidentiality leak was enabled by system-level control failure and then managed through deliberate post-leak crisis communication.

# 5. Communication Relationships and Information Flow Analysis

This section examines how communication moved across agents, channels, and interaction scopes. While the previous section focused on the timing of communication, this section focuses on the structure of communication: who was active, where they communicated, and whether information flow was concentrated or distributed.

The purpose is to identify whether TenantThread’s communication pattern looked like a controlled internal process or a broader, more fragmented multi-agent communication environment. To support visual variety, this section uses composition charts, relationship matrices, flow diagrams, and network-based views rather than relying on a single chart type.

## 5.1 Agent Communication Composition

This visual examines each agent’s communication composition across public, private, and internal scopes. Instead of only comparing total message counts, the proportional view shows whether an agent was mainly active in internal discussion, one-on-one coordination, or public-facing communication.

```{r}
agent_scope_composition <- communications_integrated %>%
  count(agent_label, communication_scope) %>%
  group_by(agent_label) %>%
  mutate(
    total_messages = sum(n),
    share = n / total_messages
  ) %>%
  ungroup()

ggplot(
  agent_scope_composition,
  aes(
    x = reorder(agent_label, total_messages),
    y = share,
    fill = communication_scope
  )
) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Agent Communication Composition by Scope",
    subtitle = "Proportion of each agent's communications across internal, private, and public scopes",
    x = "Agent",
    y = "Share of Communications",
    fill = "Communication Scope"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

This chart shows that all agents communicate mainly through internal channels, but their scope composition differs. PR-Intern-Agent and Intern-Agent have the largest public-facing shares, while Judge-Agent has no public communication and is mostly internal/private. Legal, PR, Social Manager, and Platform Trust combine internal discussion with private coordination, suggesting that key operational agents worked across multiple communication spaces.

## 5.2 Agent-Channel Relationship

This visual examines how agents used different communication channels. Unlike the previous proportional composition chart, this matrix-style bubble chart highlights the relationship between agents and channels by showing where communication activity was concentrated.

```{r}
agent_channel_matrix <- communications_integrated %>%
  count(agent_label, channel, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(agent_total = sum(message_count)) %>%
  ungroup()

ggplot(
  agent_channel_matrix,
  aes(
    x = channel,
    y = reorder(agent_label, agent_total),
    size = message_count,
    fill = communication_scope
  )
) +
  geom_point(
    shape = 21,
    alpha = 0.75,
    colour = "grey30"
  ) +
  scale_size_continuous(
    range = c(3, 14)
  ) +
  labs(
    title = "Agent-Channel Communication Matrix",
    subtitle = "Bubble size represents message volume; fill shows communication scope",
    x = "Communication Channel",
    y = "Agent",
    size = "Message Count",
    fill = "Communication Scope"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

The matrix shows that `comms_huddle` and `one_on_one_chat` are the dominant channels for most agents, while public-facing channels are used by fewer agents. Legal, Social Manager, and Platform Trust have relatively large bubbles across multiple channels, suggesting broader involvement in communication flow. PR-Intern and Intern show visible activity in `official_post` or `personal_post`, which makes them important for later public-channel analysis.

## 5.3 Reply-Based Message Flow between Agents

This visual uses reply relationships to approximate agent-to-agent communication flow. If one message responds to another message, the original message sender is treated as the source agent, while the replying agent is treated as the target agent.

```{r}
message_lookup <- communications_integrated %>%
  select(
    responding_to = message_id,
    source_agent = agent_label
  )

reply_flow <- communications_integrated %>%
  filter(!is.na(responding_to)) %>%
  left_join(
    message_lookup,
    by = "responding_to"
  ) %>%
  filter(!is.na(source_agent)) %>%
  mutate(
    target_agent = agent_label
  )

agent_flow <- reply_flow %>%
  count(
    source_agent,
    target_agent,
    name = "flow_count"
  ) %>%
  filter(source_agent != target_agent) %>%
  arrange(desc(flow_count)) %>%
  slice_max(flow_count, n = 20) %>%
  mutate(
    flow_id = row_number()
  )

flow_long <- agent_flow %>%
  select(
    flow_id,
    source_agent,
    target_agent,
    flow_count
  ) %>%
  pivot_longer(
    cols = c(source_agent, target_agent),
    names_to = "flow_side",
    values_to = "agent"
  ) %>%
  mutate(
    flow_side = recode(
      flow_side,
      source_agent = "Source Agent",
      target_agent = "Replying Agent"
    ),
    flow_side = factor(
      flow_side,
      levels = c("Source Agent", "Replying Agent")
    )
  )

ggplot(
  flow_long,
  aes(
    x = flow_side,
    y = agent,
    group = flow_id
  )
) +
  geom_line(
    aes(
      linewidth = flow_count,
      colour = flow_count
    ),
    alpha = 0.55
  ) +
  geom_point(
    aes(
      size = flow_count,
      fill = flow_count
    ),
    shape = 21,
    colour = "grey35",
    alpha = 0.9
  ) +
  scale_linewidth_continuous(
    range = c(0.4, 3),
    guide = "none"
  ) +
  scale_size_continuous(
    range = c(3, 10),
    guide = "none"
  ) +
  scale_colour_gradient(
    low = "lightblue",
    high = "darkblue",
    name = "Reply Count"
  ) +
  scale_fill_gradient(
    low = "lightblue",
    high = "darkblue",
    guide = "none"
  ) +
  labs(
    title = "Reply-Based Message Flow between Agents",
    subtitle = "Darker and thicker links indicate stronger reply-based communication flow",
    x = "",
    y = "Agent"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

This chart shows that reply-based communication is concentrated around several core agents, especially Legal-Agent, Platform-Trust-Agent, and Social-Manager-Agent. The thickest and darkest links indicate stronger reply flows between these agents, suggesting that they formed the main communication backbone. PR-Agent and PR-Intern-Agent also appear in the reply structure, but with less dominant flow intensity. This helps identify which agents should be examined more closely when tracing sensitive information movement.

## 5.4 Reply-Based Agent Relationship Network

This network visualises reply-based relationships between agents. Each node represents an agent, and each edge represents a reply relationship. Larger nodes indicate agents with more communication records, while thicker edges indicate stronger reply-based interaction between agents.

```{r}
#| fig-height: 8
#| fig-width: 11
library(igraph)
library(ggraph)

message_lookup <- communications_integrated %>%
  select(
    responding_to = message_id,
    source_agent = agent_label,
    source_role = agent_role
  )

reply_edges <- communications_integrated %>%
  filter(!is.na(responding_to)) %>%
  left_join(
    message_lookup,
    by = "responding_to"
  ) %>%
  filter(!is.na(source_agent)) %>%
  mutate(
    target_agent = agent_label,
    target_role = agent_role
  ) %>%
  filter(source_agent != target_agent) %>%
  count(
    source_agent,
    target_agent,
    name = "reply_count"
  )

agent_nodes <- communications_integrated %>%
  count(
    agent_label,
    agent_role,
    name = "message_count"
  ) %>%
  rename(
    name = agent_label,
    role = agent_role
  )

agent_network <- graph_from_data_frame(
  d = reply_edges,
  vertices = agent_nodes,
  directed = TRUE
)

ggraph(agent_network, layout = "fr") +
  geom_edge_link(
    aes(width = reply_count),
    alpha = 0.35,
    colour = "grey35",
    arrow = arrow(length = unit(2.2, "mm")),
    end_cap = circle(3, "mm")
  ) +
  geom_node_point(
    aes(
      size = message_count,
      colour = role
    ),
    alpha = 0.9
  ) +
  geom_node_text(
    aes(label = name),
    repel = TRUE,
    size = 3.4
  ) +
  scale_edge_width(
    range = c(0.3, 2.2),
    name = "Reply Count"
  ) +
  scale_size_continuous(
    range = c(5, 11),
    name = "Message Count"
  ) +
  labs(
    title = "Reply-Based Agent Relationship Network",
    subtitle = "Node size represents message volume; edge thickness represents reply frequency",
    colour = "Agent Role"
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    plot.margin = margin(10, 20, 10, 20)
  ) +
  guides(
    colour = guide_legend(
      title = "Agent Role",
      nrow = 2,
      override.aes = list(size = 4)
    ),
    size = guide_legend(
      title = "Message Count",
      nrow = 1
    ),
    edge_width = guide_legend(
      title = "Reply Count",
      nrow = 1
    )
  )
```

This network shows that Legal-Agent, Platform-Trust-Agent, Social-Manager-Agent, and PR-Agent form the core reply-based communication structure. The largest nodes and thicker links suggest that these agents were central to coordination and information exchange. PR-Intern-Agent and Intern-Agent are less central overall, but they remain connected to the main network, which matters because they also have public-facing communication roles. This structure suggests that sensitive information could move from core internal agents toward public-facing actors through existing reply relationships.

## 5.5 Communication Boundary Bridge Map

This visual identifies which agents operated across internal, private, and public communication spaces. Agents with both high internal activity and visible public-facing activity are especially important for later leak-tracing analysis because they sit closer to the boundary between internal coordination and public communication.

```{r}
bridge_data <- communications_integrated %>%
  count(agent_label, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(
    total_messages = sum(message_count),
    share = message_count / total_messages
  ) %>%
  ungroup() %>%
  select(agent_label, communication_scope, share, total_messages) %>%
  pivot_wider(
    names_from = communication_scope,
    values_from = share,
    values_fill = 0
  ) %>%
  rename(
    internal_share = Internal,
    private_share = `Private / One-on-One`,
    public_share = Public
  ) %>%
  mutate(
    bridge_score = private_share + public_share,

    label_x = case_when(
      agent_label == "Platform-Trust-Agent" ~ internal_share + 0.040,
      agent_label == "Judge-Agent" ~ internal_share - 0.050,
      agent_label == "Legal-Agent" ~ internal_share - 0.035,
      agent_label == "PR-Agent" ~ internal_share + 0.030,
      agent_label == "Social-Manager-Agent" ~ internal_share + 0.005,
      agent_label == "PR-Intern-Agent" ~ internal_share - 0.005,
      agent_label == "Intern-Agent" ~ internal_share + 0.005,
      TRUE ~ internal_share
    ),

    label_y = case_when(
      agent_label == "Platform-Trust-Agent" ~ public_share - 0.010,
      agent_label == "Judge-Agent" ~ public_share + 0.018,
      agent_label == "Legal-Agent" ~ public_share + 0.020,
      agent_label == "PR-Agent" ~ public_share + 0.030,
      agent_label == "Social-Manager-Agent" ~ public_share + 0.018,
      agent_label == "PR-Intern-Agent" ~ public_share - 0.020,
      agent_label == "Intern-Agent" ~ public_share + 0.020,
      TRUE ~ public_share + 0.015
    )
  )

ggplot(
  bridge_data,
  aes(
    x = internal_share,
    y = public_share
  )
) +
  geom_point(
    aes(
      size = total_messages,
      fill = bridge_score
    ),
    shape = 21,
    colour = "grey35",
    alpha = 0.38,
    stroke = 0.5
  ) +

  geom_segment(
    aes(
      x = internal_share,
      y = public_share,
      xend = label_x,
      yend = label_y
    ),
    colour = "grey55",
    linewidth = 0.3,
    alpha = 0.7
  ) +

  geom_label(
    aes(
      x = label_x,
      y = label_y,
      label = agent_label
    ),
    size = 3,
    label.size = 0.15,
    label.padding = unit(0.12, "lines"),
    fill = alpha("white", 0.82)
  ) +

  scale_x_continuous(
    labels = scales::percent_format(),
    breaks = seq(0.45, 0.85, 0.05)
  ) +

  scale_y_continuous(
    labels = scales::percent_format(),
    breaks = seq(0.00, 0.25, 0.05)
  ) +

  coord_cartesian(
    xlim = c(0.45, 0.85),
    ylim = c(-0.015, 0.26),
    clip = "off"
  ) +

  scale_size_continuous(
    range = c(3, 12)
  ) +

  scale_fill_gradient(
    low = "lightblue",
    high = "darkblue"
  ) +

  labs(
    title = "Communication Boundary Bridge Map",
    subtitle = "Zoomed view of agents' internal and public communication shares",
    x = "Internal Communication Share",
    y = "Public Communication Share",
    size = "Total Messages",
    fill = "Bridge Score"
  ) +

  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.margin = margin(10, 50, 10, 10)
  )
```

This bridge map shows that PR-Intern-Agent and Intern-Agent have the highest public communication shares, while Legal-Agent, PR-Agent, Social-Manager-Agent, and Platform-Trust-Agent are more embedded in internal communication. Platform-Trust-Agent corresponds to the larger bubble in the lower-right area, while Judge-Agent corresponds to the smaller nearby bubble. Agents positioned between internal and public communication spaces are important for later leak-tracing analysis because they may act as communication boundary bridges.

## 5.6 Merger-Related Communication Concentration

This visual focuses only on communications containing merger-related terms. The mosaic-style chart shows both the relative volume of merger-related communications by agent and how those communications are distributed across internal, private, and public scopes.

```{r}
merger_mosaic_data <- communications_integrated %>%
  filter(contains_merger == TRUE) %>%
  count(agent_label, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(
    agent_total = sum(message_count),
    scope_share = message_count / agent_total
  ) %>%
  ungroup()

agent_widths <- merger_mosaic_data %>%
  distinct(agent_label, agent_total) %>%
  arrange(desc(agent_total)) %>%
  mutate(
    width = agent_total / sum(agent_total),
    xmin = lag(cumsum(width), default = 0),
    xmax = cumsum(width)
  )

merger_mosaic_plot <- merger_mosaic_data %>%
  left_join(agent_widths, by = c("agent_label", "agent_total")) %>%
  group_by(agent_label) %>%
  arrange(communication_scope, .by_group = TRUE) %>%
  mutate(
    ymin = lag(cumsum(scope_share), default = 0),
    ymax = cumsum(scope_share),
    label_x = (xmin + xmax) / 2,
    label_y = (ymin + ymax) / 2
  ) %>%
  ungroup()

ggplot(
  merger_mosaic_plot,
  aes(
    xmin = xmin,
    xmax = xmax,
    ymin = ymin,
    ymax = ymax,
    fill = communication_scope
  )
) +
  geom_rect(
    color = "white",
    linewidth = 0.7,
    alpha = 0.85
  ) +
  geom_text(
    aes(
      x = label_x,
      y = label_y,
      label = ifelse(message_count > 3, message_count, "")
    ),
    size = 3
  ) +
  scale_x_continuous(
    breaks = agent_widths$xmin + agent_widths$width / 2,
    labels = agent_widths$agent_label,
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = c(0, 0)
  ) +
  labs(
    title = "Merger-Related Communication Concentration",
    subtitle = "Column width shows each agent's merger-related volume; height shows scope composition",
    x = "Agent",
    y = "Share within Agent",
    fill = "Communication Scope"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

This mosaic chart shows that merger-related communication is heavily concentrated among Legal-Agent and Social-Manager-Agent, which have the widest columns. For both agents, most merger-related records are internal, but there is also a visible public component. PR-Intern-Agent has a smaller total volume, but a relatively large private share. This suggests that merger-related information was not limited to one actor or one communication scope, making these agents important for later leak-tracing analysis.

## 5.7 Sensitive Signal Exposure by Agent

This visual compares how different agents were exposed to merger-related, confidentiality-related, and urgency-related communication signals. Unlike the previous visuals, which focus on structure and flow, this chart highlights risk-relevant content exposure by agent.

```{r}
sensitive_exposure <- communications_integrated %>%
  group_by(agent_label) %>%
  summarise(
    merger_related = sum(contains_merger, na.rm = TRUE),
    confidential_related = sum(contains_confidential, na.rm = TRUE),
    urgency_related = sum(contains_urgency, na.rm = TRUE),
    total_messages = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(merger_related, confidential_related, urgency_related),
    names_to = "signal_type",
    values_to = "signal_count"
  ) %>%
  mutate(
    signal_type = recode(
      signal_type,
      merger_related = "Merger-related",
      confidential_related = "Confidentiality-related",
      urgency_related = "Urgency-related"
    )
  )

agent_order <- sensitive_exposure %>%
  group_by(agent_label) %>%
  summarise(total_signal = sum(signal_count), .groups = "drop") %>%
  arrange(total_signal) %>%
  pull(agent_label)

sensitive_exposure <- sensitive_exposure %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order)
  )

ggplot(
  sensitive_exposure,
  aes(
    x = signal_count,
    y = agent_label,
    colour = signal_type
  )
) +
  geom_point(size = 3.2, alpha = 0.85) +
  geom_line(
    aes(group = agent_label),
    colour = "grey75",
    linewidth = 0.5
  ) +
  labs(
    title = "Sensitive Signal Exposure by Agent",
    subtitle = "Comparison of merger, confidentiality, and urgency-related signals",
    x = "Number of Communications",
    y = "Agent",
    colour = "Signal Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

Legal-Agent has the highest exposure across all three sensitive signals, especially merger- and urgency-related content. Social-Manager-Agent is also highly exposed to merger-related signals, while Platform-Trust-Agent stands out more on urgency-related content. Overall, sensitive information is concentrated among a few core agents rather than evenly distributed across the team.

## 5.8 Embargo-Relevant Relationship Profile

```{r}
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")

# Rebuild reply relationships for centrality measurement
reply_lookup <- communications_integrated %>%
  select(
    responding_to = message_id,
    source_agent = agent_label
  )

reply_links <- communications_integrated %>%
  filter(!is.na(responding_to)) %>%
  left_join(
    reply_lookup,
    by = "responding_to"
  ) %>%
  filter(!is.na(source_agent)) %>%
  mutate(
    target_agent = agent_label
  ) %>%
  filter(source_agent != target_agent)

reply_centrality <- bind_rows(
  reply_links %>%
    count(source_agent, name = "reply_links") %>%
    rename(agent_label = source_agent),
  
  reply_links %>%
    count(target_agent, name = "reply_links") %>%
    rename(agent_label = target_agent)
) %>%
  group_by(agent_label) %>%
  summarise(
    reply_centrality = sum(reply_links),
    .groups = "drop"
  )

# Agent-level embargo relevance profile
agent_embargo_profile <- communications_integrated %>%
  group_by(agent_label) %>%
  summarise(
    total_messages = n(),
    
    pre6_merger = sum(
      date == as_date("2046-06-05") &
        timestamp < embargo_lift_time &
        contains_merger,
      na.rm = TRUE
    ),
    
    pre6_public_merger = sum(
      date == as_date("2046-06-05") &
        timestamp < embargo_lift_time &
        is_public &
        contains_merger,
      na.rm = TRUE
    ),
    
    public_share = mean(is_public, na.rm = TRUE),
    
    sensitive_exposure = sum(
      as.integer(contains_merger) +
        as.integer(contains_confidential) +
        as.integer(contains_urgency),
      na.rm = TRUE
    ),
    
    .groups = "drop"
  ) %>%
  left_join(
    reply_centrality,
    by = "agent_label"
  ) %>%
  mutate(
    reply_centrality = replace_na(reply_centrality, 0)
  )

# Build communication boundary profile
boundary_profile <- communications_integrated %>%
  count(agent_label, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(
    total_scope_messages = sum(message_count),
    share = message_count / total_scope_messages
  ) %>%
  ungroup() %>%
  select(agent_label, communication_scope, share) %>%
  pivot_wider(
    names_from = communication_scope,
    values_from = share,
    values_fill = 0
  )

# Ensure expected communication scope columns exist
if (!"Internal" %in% names(boundary_profile)) {
  boundary_profile$Internal <- 0
}

if (!"Private / One-on-One" %in% names(boundary_profile)) {
  boundary_profile[["Private / One-on-One"]] <- 0
}

if (!"Public" %in% names(boundary_profile)) {
  boundary_profile$Public <- 0
}

boundary_profile <- boundary_profile %>%
  transmute(
    agent_label,
    internal_share = Internal,
    private_share = `Private / One-on-One`,
    public_scope_share = Public,
    boundary_bridge_score = private_share + public_scope_share
  )

agent_embargo_profile <- agent_embargo_profile %>%
  left_join(
    boundary_profile,
    by = "agent_label"
  ) %>%
  mutate(
    boundary_bridge_score = replace_na(boundary_bridge_score, 0)
  )

# Convert metrics into long format and scale within each metric
embargo_profile_long <- agent_embargo_profile %>%
  pivot_longer(
    cols = c(
      pre6_merger,
      pre6_public_merger,
      public_share,
      boundary_bridge_score,
      sensitive_exposure,
      reply_centrality
    ),
    names_to = "metric",
    values_to = "metric_value"
  ) %>%
  group_by(metric) %>%
  mutate(
    scaled_score = {
      metric_min <- min(metric_value, na.rm = TRUE)
      metric_max <- max(metric_value, na.rm = TRUE)
      
      if (is.infinite(metric_min) | is.infinite(metric_max) | metric_max == metric_min) {
        rep(0, n())
      } else {
        (metric_value - metric_min) / (metric_max - metric_min)
      }
    }
  ) %>%
  ungroup() %>%
  mutate(
    metric = recode(
      metric,
      pre6_merger = "Pre-6PM Merger Activity",
      pre6_public_merger = "Pre-6PM Public Merger",
      public_share = "Public-Facing Share",
      boundary_bridge_score = "Boundary Bridge Score",
      sensitive_exposure = "Sensitive Signal Exposure",
      reply_centrality = "Reply Centrality"
    ),
    
    metric = factor(
      metric,
      levels = c(
        "Pre-6PM Merger Activity",
        "Pre-6PM Public Merger",
        "Public-Facing Share",
        "Boundary Bridge Score",
        "Sensitive Signal Exposure",
        "Reply Centrality"
      )
    ),
    
    label = case_when(
      metric %in% c("Public-Facing Share", "Boundary Bridge Score") ~
        scales::percent(metric_value, accuracy = 1),
      TRUE ~ as.character(round(metric_value, 0))
    )
  )

agent_order <- embargo_profile_long %>%
  group_by(agent_label) %>%
  summarise(
    total_score = sum(scaled_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(total_score) %>%
  pull(agent_label)

embargo_profile_long <- embargo_profile_long %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order)
  )

ggplot(
  embargo_profile_long,
  aes(
    x = metric,
    y = agent_label,
    fill = scaled_score
  )
) +
  geom_tile(
    color = "white",
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = label),
    size = 3.1
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkred"
  ) +
  labs(
    title = "Embargo-Relevant Relationship Profile",
    subtitle = "Agent-level comparison of pre-embargo merger activity, public exposure, sensitivity, and reply centrality",
    x = "Embargo-Relevant Dimension",
    y = "Agent",
    fill = "Relative Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

Legal-Agent and Social-Manager-Agent stand out as the most embargo-relevant actors because they combine high pre-6PM merger activity with strong sensitive-signal exposure. Platform-Trust-Agent is important through reply centrality and boundary relevance, suggesting a coordination role. PR-Intern-Agent and Intern-Agent show stronger public-facing exposure, which makes them important for tracing how merger-related information could move toward public channels before the 6:00 PM embargo lift.

## 5.9 Key Findings from Communication Flow Analysis

- **Communication was concentrated around a core coordination group.**\
  The reply-based flow and network graphs show that Legal-Agent, Platform-Trust-Agent, Social-Manager-Agent, and PR-Agent formed the main communication backbone. These agents were central to information exchange before the breach, which means that merger-related risk was managed through a small group of highly connected coordinators rather than through a fully isolated legal channel.

- **Merger-related information was not fully contained within legal control.**\
  Legal-Agent and Social-Manager-Agent had the highest pre-6PM merger activity and sensitive-signal exposure, but merger-related content also appeared across other agents, scopes, and channels. This suggests that confidential merger information had moved beyond a narrow legal-review context before the official embargo lift.

- **Public-facing boundary actors created potential release pathways.**\
  PR-Intern-Agent and Intern-Agent had higher public-facing shares, while Social-Manager-Agent connected internal coordination with public-facing communication. This matters because these agents sat closer to the boundary between internal information and public release, making them important for understanding how sensitive information could move from internal discussion toward public-facing spaces.

- **The embargo-relevant profile highlights different risk roles.**\
  Legal-Agent and Social-Manager-Agent were most important for merger and sensitive-content exposure; Platform-Trust-Agent was important as a reply-centrality and coordination node; PR-Intern-Agent and Intern-Agent were important because of their public-facing exposure. These roles indicate that the breach pathway was distributed across legal, monitoring, coordination, and public-facing functions.

- **Overall, the communication structure supports an internal-leak-enabled system failure interpretation.**\
  Part 5 does not identify the original source of the Slack screenshot or prove that a senior agent deliberately leaked the merger. Instead, the graphs show the communication environment that made the breach difficult to contain: sensitive merger information circulated through core coordinators while public-facing boundary actors remained active before the embargo lift. Therefore, the network evidence should be read together with the Slack timeline from Part 4. The Slack screenshot provides the direct internal-leak evidence, while the Part 5 communication-flow visuals explain how that leak could be amplified through a fragmented multi-agent communication system.

# 6. Key Events, Decision Points, and Causal Chain

This section reconstructs the event-level pathway that led to the inappropriate release. Part 4 showed when communication and merger-related signals escalated, while Part 5 showed which agents and channels formed the main information pathways. This section connects those findings by identifying key actions, decision points, actors involved, and system elements that allowed embargoed merger information to move into public-facing communication before the official 6:00 PM embargo lift.

## 6.1 Purpose of Causal Event Analysis

The purpose of this section is to move from descriptive patterns to causal explanation. Earlier sections established that communication volume, merger-related discussion, public-facing exposure, and pressure signals all intensified before the embargo breach. However, those findings alone do not explain how the inappropriate release happened.

This section therefore reconstructs the event-level pathway that led to the pre-6:00 PM release of merger-related information. It focuses on three questions:

1.  Which key actions and decision points appeared before the embargo lift?
2.  Which agents were involved in those decisions or actions?
3.  Which system elements allowed merger-related content to move from internal or private communication into public-facing channels?

The analysis is especially focused on the period before the official 6:00 PM embargo lift on June 5. Records involving merger-related content, decision/action language, public-facing communication, and enforcement-related signals are treated as key evidence for reconstructing the causal chain.

This section does not assume that the release was either deliberate or accidental at the outset. Instead, it evaluates whether the event pattern looks like a controlled intentional leak by a specific actor, or a broader system breakdown in which sensitive information passed through multiple agents, channels, and weak control points before reaching public communication.

## 6.2 Condensed Key Event Sequence

This section condenses the communication records into a small number of causal stages. The purpose is not to repeat the full timeline analysis, but to show the main pathway from internal merger circulation to public-facing exposure before and after the embargo lift.

```{r}
#| fig-width: 11
#| fig-height: 6.5

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

causal_event_data <- communications_integrated %>%
  mutate(
    event_text = str_to_lower(content),

    causal_stage = case_when(
      contains_merger &
        !is_public &
        timestamp < breach_time ~
        "Internal merger circulation",

      contains_merger &
        str_detect(event_text, "urgent|asap|deadline|pressure|crisis|escalate|critical|risk") &
        timestamp < embargo_lift_time ~
        "Merger pressure escalation",

      contains_merger &
        str_detect(event_text, "embargo|confidential|restricted|sensitive|not public|internal only") &
        timestamp < embargo_lift_time ~
        "Confidentiality boundary present",

      str_detect(event_text, "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release") &
        timestamp < embargo_lift_time ~
        "Decision / release action",

      str_detect(event_text, "hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready") &
        timestamp < embargo_lift_time ~
        "Review / control attempt",

      str_detect(event_text, "judge|enforcement|guardrail|compliance|policy") &
        timestamp < embargo_lift_time ~
        "Enforcement reference",

      is_public &
        contains_merger &
        timestamp < embargo_lift_time ~
        "Public merger exposure before embargo lift",

      is_public &
        contains_merger &
        timestamp >= embargo_lift_time ~
        "Public merger release after embargo lift",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(causal_stage))

causal_stage_summary <- causal_event_data %>%
  group_by(causal_stage) %>%
  summarise(
    event_count = n(),
    agents_involved = n_distinct(agent_label),
    public_events = sum(is_public, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    stage_number = case_when(
      causal_stage == "Internal merger circulation" ~ 1,
      causal_stage == "Merger pressure escalation" ~ 2,
      causal_stage == "Confidentiality boundary present" ~ 3,
      causal_stage == "Decision / release action" ~ 4,
      causal_stage == "Review / control attempt" ~ 5,
      causal_stage == "Enforcement reference" ~ 6,
      causal_stage == "Public merger exposure before embargo lift" ~ 7,
      causal_stage == "Public merger release after embargo lift" ~ 8
    ),

    stage_group = case_when(
      causal_stage == "Internal merger circulation" ~ "Information Flow",
      causal_stage %in% c(
        "Merger pressure escalation",
        "Confidentiality boundary present"
      ) ~ "Pressure / Boundary",
      causal_stage %in% c(
        "Decision / release action",
        "Review / control attempt",
        "Enforcement reference"
      ) ~ "Decision / Control",
      causal_stage %in% c(
        "Public merger exposure before embargo lift",
        "Public merger release after embargo lift"
      ) ~ "Public Exposure"
    ),

    stage_label = paste0(stage_number, ". ", causal_stage),
    stage_label = str_wrap(stage_label, width = 32),
    stage_label = factor(stage_label, levels = rev(stage_label)),

    detail_label = paste0(
      event_count, " events",
      " | ", agents_involved, " agents",
      " | ", public_events, " public"
    )
  ) %>%
  arrange(stage_number)

ggplot(
  causal_stage_summary,
  aes(
    x = event_count,
    y = stage_label,
    fill = stage_group
  )
) +
  geom_col(
    width = 0.62,
    alpha = 0.85
  ) +
  geom_point(
    aes(size = agents_involved),
    shape = 21,
    fill = "white",
    colour = "grey25",
    stroke = 0.8
  ) +
  geom_text(
    aes(label = detail_label),
    hjust = -0.05,
    size = 3.4
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.35))
  ) +
  scale_fill_manual(
    values = c(
      "Information Flow" = "#4E79A7",
      "Pressure / Boundary" = "#F28E2B",
      "Decision / Control" = "#B07AA1",
      "Public Exposure" = "#E15759"
    )
  ) +
  scale_size_continuous(
    range = c(3.5, 8),
    name = "Agents involved"
  ) +
  labs(
    title = "Condensed Key Event Sequence",
    subtitle = "Causal stages from internal merger circulation to public exposure",
    x = "Number of Events",
    y = "Causal Stage",
    fill = "Stage Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 9),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

The chart summarises the breach pathway across eight causal stages. Decision/release action and review/control attempt have the largest event volumes, showing that action and control signals were both active before the embargo lift. However, public merger exposure still appeared before 6:00 PM, meaning that existing control and enforcement activity did not stop merger-related information from reaching public-facing communication.

## 6.3 Decision and Control Failure Points

This section focuses only on the key decision and control nodes that were directly relevant to the embargo failure: approval or release actions, hold or review controls, enforcement references, and public merger exposure before the official 6:00 PM embargo lift.

```{r}
#| fig-width: 11
#| fig-height: 7

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

decision_control_data <- communications_integrated %>%
  mutate(
    event_text = str_to_lower(replace_na(content, "")),
    
    breach_phase = case_when(
      date < as_date("2046-06-05") ~ "Before June 5",
      date == as_date("2046-06-05") & timestamp < breach_time ~ "June 5 Before 5PM",
      date == as_date("2046-06-05") &
        timestamp >= breach_time &
        timestamp < embargo_lift_time ~ "5–6PM Breach Window",
      TRUE ~ NA_character_
    ),
    
    approval_release = str_detect(
      event_text,
      "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release"
    ) & timestamp < embargo_lift_time,
    
    hold_review = str_detect(
      event_text,
      "hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready"
    ) & timestamp < embargo_lift_time,
    
    enforcement_reference = str_detect(
      event_text,
      "judge|enforcement|guardrail|compliance|policy"
    ) & timestamp < embargo_lift_time,
    
    public_merger_exposure = is_public &
      contains_merger &
      timestamp < embargo_lift_time
  ) %>%
  filter(!is.na(breach_phase)) %>%
  mutate(
    breach_phase = factor(
      breach_phase,
      levels = c(
        "Before June 5",
        "June 5 Before 5PM",
        "5–6PM Breach Window"
      )
    )
  )

decision_control_summary <- decision_control_data %>%
  summarise(
    `Approval / release action` = sum(approval_release, na.rm = TRUE),
    `Hold / review control` = sum(hold_review, na.rm = TRUE),
    `Enforcement reference` = sum(enforcement_reference, na.rm = TRUE),
    `Public merger exposure before 6PM` = sum(public_merger_exposure, na.rm = TRUE),
    .by = breach_phase
  ) %>%
  pivot_longer(
    cols = -breach_phase,
    names_to = "control_node",
    values_to = "event_count"
  ) %>%
  mutate(
    control_node = factor(
      control_node,
      levels = c(
        "Approval / release action",
        "Hold / review control",
        "Enforcement reference",
        "Public merger exposure before 6PM"
      )
    )
  )

ggplot(
  decision_control_summary,
  aes(
    x = breach_phase,
    y = event_count,
    fill = control_node
  )
) +
  geom_col(
    width = 0.65,
    alpha = 0.85
  ) +
  geom_text(
    aes(label = event_count),
    vjust = -0.4,
    size = 3.5
  ) +
  facet_wrap(
    ~ control_node,
    scales = "free_y",
    ncol = 2
  ) +
  scale_fill_manual(
    values = c(
      "Approval / release action" = "#4E79A7",
      "Hold / review control" = "#F28E2B",
      "Enforcement reference" = "#B07AA1",
      "Public merger exposure before 6PM" = "#E15759"
    )
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title = "Decision and Control Failure Points",
    subtitle = "Key approval, hold, enforcement, and public-exposure nodes across the pre-embargo phases",
    x = "Pre-Embargo Phase",
    y = "Event Count",
    fill = "Control Node"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 20, hjust = 1),
    strip.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

The chart shows that approval/release actions and hold/review controls were already active before the breach window, especially on June 5 before 5PM. However, public merger exposure also appeared before 6PM, with 9 cases before 5PM and 7 more during the 5–6PM breach window. This indicates that the problem was not the absence of decision or control activity, but the failure of those controls to stop merger-related information from reaching public-facing communication.

## 6.4 Embargo Enforcement Failure Pathway

This section summarises the causal pathway that allowed embargoed merger information to move from internal communication into public-facing exposure before the official embargo lift. The flow direction shows the causal sequence, while bubble size gives a visual indication of the evidence volume at each stage.

```{r}
#| fig-width: 12
#| fig-height: 5.8

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")

event_text <- str_to_lower(replace_na(communications_integrated$content, ""))

pathway_nodes <- tibble(
  step = 1:5,
  x = c(1, 2.25, 3.5, 4.75, 6),
  y = 1,
  stage = c(
    "Internal merger\ncirculation",
    "Pressure +\nconfidentiality",
    "Decision / review /\nenforcement activity",
    "Public-facing\nboundary",
    "Pre-6PM public\nmerger exposure"
  ),
  stage_group = c(
    "Information Flow",
    "Pressure / Sensitivity",
    "Decision / Control",
    "Boundary Exposure",
    "Embargo Failure"
  ),
  value_type = c(
    "events",
    "events",
    "events",
    "agents",
    "events"
  ),
  value = c(
    sum(
      communications_integrated$contains_merger &
        !communications_integrated$is_public &
        communications_integrated$timestamp < embargo_lift_time,
      na.rm = TRUE
    ),
    
    sum(
      communications_integrated$contains_merger &
        (
          str_detect(event_text, "urgent|asap|deadline|pressure|crisis|escalate|critical|risk") |
            str_detect(event_text, "embargo|confidential|restricted|sensitive|not public|internal only")
        ) &
        communications_integrated$timestamp < embargo_lift_time,
      na.rm = TRUE
    ),
    
    sum(
      (
        str_detect(event_text, "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release") |
          str_detect(event_text, "hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready") |
          str_detect(event_text, "judge|enforcement|guardrail|compliance|policy")
      ) &
        communications_integrated$timestamp < embargo_lift_time,
      na.rm = TRUE
    ),
    
    n_distinct(
      communications_integrated$agent_label[
        communications_integrated$is_public &
          communications_integrated$timestamp < embargo_lift_time
      ]
    ),
    
    sum(
      communications_integrated$is_public &
        communications_integrated$contains_merger &
        communications_integrated$timestamp < embargo_lift_time,
      na.rm = TRUE
    )
  )
) %>%
  mutate(
    display_size = scales::rescale(
      sqrt(value),
      to = c(10, 24)
    ),
    value_label = paste0(value, " ", value_type)
  )

pathway_edges <- pathway_nodes %>%
  arrange(step) %>%
  mutate(
    xend = lead(x),
    yend = lead(y)
  ) %>%
  filter(!is.na(xend)) %>%
  mutate(
    x_start = x + 0.38,
    x_end = xend - 0.38
  )

ggplot() +
  geom_segment(
    data = pathway_edges,
    aes(
      x = x_start,
      xend = x_end,
      y = y,
      yend = yend
    ),
    arrow = arrow(
      length = unit(0.25, "cm"),
      type = "closed"
    ),
    linewidth = 0.9,
    colour = "grey35"
  ) +
  geom_point(
    data = pathway_nodes,
    aes(
      x = x,
      y = y,
      fill = stage_group
    ),
    size = pathway_nodes$display_size,
    shape = 21,
    colour = "grey20",
    stroke = 0.9,
    alpha = 0.9
  ) +
  geom_text(
    data = pathway_nodes,
    aes(
      x = x,
      y = 1.42,
      label = value_label
    ),
    size = 3.7,
    fontface = "bold"
  ) +
  geom_text(
    data = pathway_nodes,
    aes(
      x = x,
      y = 0.45,
      label = stage
    ),
    size = 3.6,
    lineheight = 0.95
  ) +
  geom_text(
    data = pathway_nodes,
    aes(
      x = x,
      y = 0.12,
      label = paste0("Step ", step)
    ),
    size = 3.2,
    colour = "grey35"
  ) +
  scale_fill_manual(
    values = c(
      "Information Flow" = "#4E79A7",
      "Pressure / Sensitivity" = "#F28E2B",
      "Decision / Control" = "#B07AA1",
      "Boundary Exposure" = "#59A14F",
      "Embargo Failure" = "#E15759"
    )
  ) +
  scale_x_continuous(
    limits = c(0.45, 6.55),
    breaks = NULL
  ) +
  scale_y_continuous(
    limits = c(0, 1.65),
    breaks = NULL
  ) +
  labs(
    title = "Embargo Enforcement Failure Pathway",
    subtitle = "Flow direction shows causal sequence; bubble size indicates relative evidence volume",
    x = NULL,
    y = NULL,
    fill = "Pathway Stage"
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    plot.margin = margin(10, 20, 10, 20)
  )
```

This pathway summarises how the breach moved from internal merger circulation to public-facing exposure before the embargo lift. The key insight is that decision, review, and enforcement activity existed before the release, but those controls did not prevent merger-related information from crossing the public-facing boundary.

## 6.5 Evidence Interpretation Matrix

```{r}
evidence_matrix <- tibble::tribble(
  ~evidence, ~interpretation, ~strength, ~label,
  "Employee Slack screenshot appeared on Said-it/r/PropTech",
  "Internal Leak", 3, "Strong",
  "Employee Slack screenshot appeared on Said-it/r/PropTech",
  "System Breakdown", 2, "Moderate",
  "Employee Slack screenshot appeared on Said-it/r/PropTech",
  "Senior-Orchestrated Leak", 0, "Not proven",
  
  "Legal described the information perimeter as compromised",
  "Internal Leak", 3, "Strong",
  "Legal described the information perimeter as compromised",
  "System Breakdown", 3, "Strong",
  "Legal described the information perimeter as compromised",
  "Senior-Orchestrated Leak", 0, "Not proven",
  
  "Legal linked the Slack leak to Section 8.1 confidentiality obligation",
  "Internal Leak", 3, "Strong",
  "Legal linked the Slack leak to Section 8.1 confidentiality obligation",
  "System Breakdown", 2, "Moderate",
  "Legal linked the Slack leak to Section 8.1 confidentiality obligation",
  "Senior-Orchestrated Leak", 0, "Not proven",
  
  "Public merger exposure appeared before the 6PM embargo lift",
  "Internal Leak", 3, "Strong",
  "Public merger exposure appeared before the 6PM embargo lift",
  "System Breakdown", 3, "Strong",
  "Public merger exposure appeared before the 6PM embargo lift",
  "Senior-Orchestrated Leak", 1, "Weak",
  
  "Multiple agents and channels were involved before the breach",
  "Internal Leak", 2, "Moderate",
  "Multiple agents and channels were involved before the breach",
  "System Breakdown", 3, "Strong",
  "Multiple agents and channels were involved before the breach",
  "Senior-Orchestrated Leak", 1, "Weak",
  
  "Approval, hold, review, and enforcement controls were active but ineffective",
  "Internal Leak", 2, "Moderate",
  "Approval, hold, review, and enforcement controls were active but ineffective",
  "System Breakdown", 3, "Strong",
  "Approval, hold, review, and enforcement controls were active but ineffective",
  "Senior-Orchestrated Leak", 1, "Weak",
  
  "No clear senior actor was identified as the original leak source",
  "Internal Leak", 1, "Weak",
  "No clear senior actor was identified as the original leak source",
  "System Breakdown", 2, "Moderate",
  "No clear senior actor was identified as the original leak source",
  "Senior-Orchestrated Leak", 0, "Not proven",
  
  "Post-leak accelerated announcement and narrative management",
  "Internal Leak", 2, "Moderate",
  "Post-leak accelerated announcement and narrative management",
  "System Breakdown", 3, "Strong",
  "Post-leak accelerated announcement and narrative management",
  "Senior-Orchestrated Leak", 2, "Moderate"
) %>%
  mutate(
    interpretation = factor(
      interpretation,
      levels = c(
        "Internal Leak",
        "System Breakdown",
        "Senior-Orchestrated Leak"
      )
    ),
    evidence = factor(
      evidence,
      levels = rev(unique(evidence))
    )
  )

ggplot(evidence_matrix, aes(x = interpretation, y = evidence, fill = strength)) +
  geom_tile(color = "white", linewidth = 0.7) +
  geom_text(aes(label = label), size = 3.2) +
  scale_fill_gradient(
    low = "yellow",
    high = "red",
    breaks = c(0, 1, 2, 3),
    labels = c("Not proven", "Weak", "Moderate", "Strong")
  ) +
  labs(
    title = "Evidence Interpretation Matrix",
    subtitle = "Evidence supports internal leakage and system control failure, but not a proven senior-orchestrated leak",
    x = NULL,
    y = NULL,
    fill = "Evidence strength"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 8),
    axis.text.x = element_text(size = 7, face = "bold"),
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )
```

This evidence interpretation matrix shows that the strongest support is for **internal leakage** and **system breakdown**, rather than a proven senior-orchestrated leak. The Slack screenshot, the compromised information perimeter, and the Section 8.1 confidentiality issue all provide strong evidence that internal merger information reached the public domain before the authorised release. At the same time, the high scores under system breakdown show that the leak became serious because multiple agents, channels, and ineffective controls failed to contain the information under pressure. In contrast, the evidence for a senior-orchestrated leak remains weak or not proven, since no clear senior actor is identified as the original source. Overall, the matrix supports a mixed interpretation: the breach was an internal confidentiality leak enabled by system-level control failure, followed by deliberate post-leak crisis management.

## 6.6 Key Findings from Causal Chain Analysis

- **The inappropriate release followed a multi-stage pathway rather than a single isolated event.**\
  Merger information first circulated internally, then became linked with pressure, confidentiality, decision/action, review/control, and enforcement-related activity before reaching public-facing exposure. This suggests that the breach developed through a sequence of escalating conditions rather than through one simple communication mistake.

- **The Slack screenshot provides direct evidence of internal leakage within the causal chain.**\
  The causal pathway should not be interpreted as only external speculation or ordinary system pressure. The dataset records an employee Slack message stating that "the merger saves all our jobs" appearing on Said-it/r/PropTech, while Legal treated internal Slack screenshots reaching Said-it within about 20 minutes as a compromised information perimeter. This places a direct internal-leak event inside the causal chain.

- **Control activity was present, but it was not effective.**\
  The decision/control analysis shows large volumes of approval/release actions and hold/review controls before the embargo lift. This means the problem was not a complete absence of review or enforcement, but a failure of those controls to prevent internal information from reaching public-facing spaces.

- **The breach became visible before the official 6:00 PM embargo lift.**\
  Public merger exposure appeared before 6PM, including cases before 5PM and during the 5–6PM breach window. This confirms that the embargo boundary had already been crossed before the authorised release time.

- **The causal pathway involved multiple system layers.**\
  The failure pathway shows internal merger circulation, pressure/confidentiality signals, decision/review/enforcement activity, public-facing boundary agents, and final pre-6PM public exposure. This indicates that the release was produced by interaction between an internal confidentiality breach and a weak multi-agent communication-control system.

- **The evidence supports internal leak plus system breakdown, rather than a clean binary conclusion.**\
  The evidence matrix should not be read as proving either a purely accidental system breakdown or a fully deliberate senior-orchestrated leak. The Slack screenshot strongly supports the internal-leak interpretation, while the wider pattern of multiple agents, multiple channels, active but ineffective controls, and rising crisis pressure supports the system-breakdown interpretation. However, the available evidence does not clearly identify a single senior actor who deliberately planned the original leak. Therefore, the strongest causal conclusion is that the merger exposure resulted from an internal confidentiality leak enabled by system-level control failure, followed by deliberate post-leak crisis management.

# 7. Typical vs Release Behavior Comparison

## 7.1 Purpose of Typical Behavior Comparison

The purpose of this section is to determine whether the behavior that led to the inappropriate release was consistent with prior system behavior or represented a meaningful deviation. Earlier sections reconstructed the breach timeline, communication relationships, and causal pathway. This section now compares the release-window behavior against the system’s typical behavior before June 5.

The analysis focuses on three questions:

1.  What did normal communication behavior look like before the breach day?
2.  How did the June 5 release-window behavior differ from that baseline?
3.  Was pre-embargo public merger exposure a continuation of prior behavior, or a new and unusual pattern?

This comparison is important because a large number of leaked messages does not automatically prove deliberate intent. If the release-window behavior differs sharply from prior patterns across agents, channels, scopes, and message types, the evidence is more consistent with a system behavior shift under pressure. If the behavior had been common before, it would suggest that the breach was part of a recurring uncontrolled pattern rather than a new failure.

## 7.2 Typical Communication Baseline Before June 5

This section establishes the typical communication baseline before the breach day. The purpose is to identify how agents normally behaved before June 5 across communication scope, merger-related content, public exposure, and decision/control language.

```{r}

baseline_data <- communications_integrated %>%
  filter(date < as_date("2046-06-05")) %>%
  mutate(
    event_text = str_to_lower(replace_na(content, "")),
    
    decision_control_signal = str_detect(
      event_text,
      "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release|hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready|judge|enforcement|guardrail|compliance|policy"
    )
  )

baseline_agent_profile <- baseline_data %>%
  group_by(agent_label) %>%
  summarise(
    total_messages = n(),
    
    internal_share = mean(communication_scope == "Internal", na.rm = TRUE),
    private_share = mean(communication_scope == "Private / One-on-One", na.rm = TRUE),
    public_share = mean(communication_scope == "Public", na.rm = TRUE),
    
    merger_share = mean(contains_merger, na.rm = TRUE),
    public_merger_share = mean(is_public & contains_merger, na.rm = TRUE),
    decision_control_share = mean(decision_control_signal, na.rm = TRUE),
    
    .groups = "drop"
  )

baseline_heatmap_data <- baseline_agent_profile %>%
  pivot_longer(
    cols = c(
      internal_share,
      private_share,
      public_share,
      merger_share,
      public_merger_share,
      decision_control_share
    ),
    names_to = "baseline_metric",
    values_to = "metric_value"
  ) %>%
  mutate(
    baseline_metric = recode(
      baseline_metric,
      internal_share = "Internal Communication",
      private_share = "Private / One-on-One",
      public_share = "Public Communication",
      merger_share = "Merger-Related Content",
      public_merger_share = "Public + Merger Content",
      decision_control_share = "Decision / Control Language"
    ),
    
    baseline_metric = factor(
      baseline_metric,
      levels = c(
        "Internal Communication",
        "Private / One-on-One",
        "Public Communication",
        "Merger-Related Content",
        "Public + Merger Content",
        "Decision / Control Language"
      )
    ),
    
    label = scales::percent(metric_value, accuracy = 1)
  )

agent_order_baseline <- baseline_heatmap_data %>%
  group_by(agent_label) %>%
  summarise(
    baseline_intensity = sum(metric_value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(baseline_intensity) %>%
  pull(agent_label)

baseline_heatmap_data <- baseline_heatmap_data %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order_baseline)
  )

ggplot(
  baseline_heatmap_data,
  aes(
    x = baseline_metric,
    y = agent_label,
    fill = metric_value
  )
) +
  geom_tile(
    colour = "white",
    linewidth = 0.8
  ) +
  geom_text(
    aes(label = label),
    size = 3.2
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkred",
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Typical Communication Baseline Before June 5",
    subtitle = "Agent-level baseline across communication scope, merger content, and control language",
    x = "Baseline Behavior Metric",
    y = "Agent",
    fill = "Share of Messages"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This baseline shows that before June 5, most agents mainly operated in internal or private communication spaces, while public communication was concentrated among PR-Intern-Agent and Intern-Agent. Public + merger content was 0% across all agents, which is the most important baseline finding. This means that pre-embargo public merger exposure was not part of normal prior behavior, making the June 5 release-window behavior a clear deviation from the typical communication pattern.

## 7.2 Typical Communication Baseline Before June 5

This section establishes the typical communication baseline before the breach day. The purpose is to identify how agents normally communicated before June 5 across communication scope, channel, and message type. This baseline provides the reference point for comparing whether the June 5 release-window behavior was a normal continuation of prior behavior or a meaningful deviation.

```{r}
#| fig-width: 12
#| fig-height: 8

baseline_data <- communications_integrated %>%
  filter(date < as_date("2046-06-05"))

baseline_scope <- baseline_data %>%
  count(
    agent_label,
    communication_scope,
    name = "message_count"
  ) %>%
  group_by(agent_label) %>%
  mutate(
    share = message_count / sum(message_count),
    baseline_dimension = "Communication Scope",
    category = communication_scope
  ) %>%
  ungroup()

baseline_channel <- baseline_data %>%
  count(
    agent_label,
    channel,
    name = "message_count"
  ) %>%
  group_by(agent_label) %>%
  mutate(
    share = message_count / sum(message_count),
    baseline_dimension = "Channel",
    category = channel
  ) %>%
  ungroup()

baseline_message_type <- baseline_data %>%
  count(
    agent_label,
    message_type,
    name = "message_count"
  ) %>%
  group_by(agent_label) %>%
  mutate(
    share = message_count / sum(message_count),
    baseline_dimension = "Message Type",
    category = message_type
  ) %>%
  ungroup()

baseline_structure <- bind_rows(
  baseline_scope,
  baseline_channel,
  baseline_message_type
) %>%
  mutate(
    baseline_dimension = factor(
      baseline_dimension,
      levels = c(
        "Communication Scope",
        "Channel",
        "Message Type"
      )
    ),
    category = str_wrap(category, width = 18),
    label = ifelse(
      share >= 0.12,
      scales::percent(share, accuracy = 1),
      ""
    )
  )

agent_order_baseline <- baseline_data %>%
  count(agent_label, name = "total_messages") %>%
  arrange(total_messages) %>%
  pull(agent_label)

baseline_structure <- baseline_structure %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order_baseline)
  )

ggplot(
  baseline_structure,
  aes(
    x = agent_label,
    y = share,
    fill = category
  )
) +
  geom_col(
    width = 0.72,
    colour = "white",
    linewidth = 0.3
  ) +
  geom_text(
    aes(label = label),
    position = position_stack(vjust = 0.5),
    size = 2.7,
    colour = "black"
  ) +
  facet_wrap(
    ~ baseline_dimension,
    ncol = 1,
    scales = "free_y"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Typical Communication Baseline Before June 5",
    subtitle = "Pre-breach agent behavior across scope, channel, and message type",
    x = "Agent",
    y = "Share of Messages",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(size = 11, face = "bold"),
    panel.grid.major.x = element_blank(),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This baseline shows that before June 5, agents had clearly different normal communication patterns. PR-Intern-Agent and Intern-Agent were the most public-facing, while Legal-Agent, Platform-Trust-Agent, Social-Manager-Agent, and Judge-Agent were mainly internal or private. The channel and message-type panels also show that core coordination agents relied more on `comms_huddle`, `one_on_one_chat`, and broadcast-style messages, while public-post behavior was concentrated in fewer boundary-facing agents. This baseline is important because it shows what “normal” looked like before comparing it with the June 5 release-window behavior.

## 7.3 Release-Window Behavior Profile

This section compares each agent’s typical pre-June 5 behavior with its behavior during the 5–6PM breach window. The comparison focuses on four release-relevant indicators: public communication, merger-related content, public merger exposure, and decision/control language.

```{r}
#| fig-width: 12
#| fig-height: 7.5

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

behavior_comparison_data <- communications_integrated %>%
  mutate(
    event_text = str_to_lower(replace_na(content, "")),
    
    period = case_when(
      date < as_date("2046-06-05") ~ "Baseline before June 5",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5–6PM breach window",
      TRUE ~ NA_character_
    ),
    
    decision_control_signal = str_detect(
      event_text,
      "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release|hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready|judge|enforcement|guardrail|compliance|policy"
    )
  ) %>%
  filter(!is.na(period))

agent_behavior_profile <- behavior_comparison_data %>%
  group_by(agent_label, period) %>%
  summarise(
    total_messages = n(),
    public_share = mean(is_public, na.rm = TRUE),
    merger_share = mean(contains_merger, na.rm = TRUE),
    public_merger_share = mean(is_public & contains_merger, na.rm = TRUE),
    decision_control_share = mean(decision_control_signal, na.rm = TRUE),
    .groups = "drop"
  )

all_agents <- sort(unique(communications_integrated$agent_label))

agent_behavior_profile <- tidyr::complete(
  agent_behavior_profile,
  agent_label = all_agents,
  period = c("Baseline before June 5", "5–6PM breach window"),
  fill = list(
    total_messages = 0,
    public_share = 0,
    merger_share = 0,
    public_merger_share = 0,
    decision_control_share = 0
  )
)

behavior_long <- agent_behavior_profile %>%
  pivot_longer(
    cols = c(
      public_share,
      merger_share,
      public_merger_share,
      decision_control_share
    ),
    names_to = "behavior_metric",
    values_to = "share"
  ) %>%
  mutate(
    behavior_metric = recode(
      behavior_metric,
      public_share = "Public Communication",
      merger_share = "Merger-Related Content",
      public_merger_share = "Public + Merger Exposure",
      decision_control_share = "Decision / Control Language"
    ),
    behavior_metric = factor(
      behavior_metric,
      levels = c(
        "Public Communication",
        "Merger-Related Content",
        "Public + Merger Exposure",
        "Decision / Control Language"
      )
    ),
    period = factor(
      period,
      levels = c(
        "Baseline before June 5",
        "5–6PM breach window"
      )
    )
  )

behavior_wide <- behavior_long %>%
  select(agent_label, behavior_metric, period, share) %>%
  pivot_wider(
    names_from = period,
    values_from = share
  ) %>%
  mutate(
    shift = `5–6PM breach window` - `Baseline before June 5`
  )

agent_order_release <- behavior_wide %>%
  group_by(agent_label) %>%
  summarise(
    total_abs_shift = sum(abs(shift), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(total_abs_shift) %>%
  pull(agent_label)

behavior_wide <- behavior_wide %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order_release)
  )

ggplot(
  behavior_wide,
  aes(y = agent_label)
) +
  geom_segment(
    aes(
      x = `Baseline before June 5`,
      xend = `5–6PM breach window`,
      yend = agent_label
    ),
    linewidth = 0.8,
    colour = "grey65"
  ) +
  geom_point(
    aes(
      x = `Baseline before June 5`,
      colour = "Baseline before June 5"
    ),
    size = 3
  ) +
  geom_point(
    aes(
      x = `5–6PM breach window`,
      colour = "5–6PM breach window"
    ),
    size = 3.4
  ) +
  facet_wrap(
    ~ behavior_metric,
    ncol = 2,
    scales = "free_x"
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0.03, 0.12))
  ) +
  scale_colour_manual(
    values = c(
      "Baseline before June 5" = "#4E79A7",
      "5–6PM breach window" = "#E15759"
    )
  ) +
  labs(
    title = "Release-Window Behavior Shift from Baseline",
    subtitle = "Connected dots compare each agent's baseline behavior with the 5–6PM breach-window behavior",
    x = "Share of Messages",
    y = "Agent",
    colour = "Period"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 9),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This chart shows a clear shift from baseline behavior during the 5–6PM breach window. The strongest deviation appears in merger-related content: Legal-Agent and Social-Manager-Agent moved from low baseline merger shares to much higher release-window shares. Public + merger exposure also appears only in the breach window for Legal-Agent and Social-Manager-Agent, making it a new behavior rather than a normal pattern. Decision/control language remains high for several core agents, which suggests that the release-window behavior changed under active control pressure rather than in the absence of controls.

## 7.4 Behavioral Deviation Score by Agent

This section aggregates the release-window behavioral shifts into an agent-level deviation profile. Each radar panel shows how much an agent’s 5–6PM breach-window behavior deviated from its pre-June 5 baseline across four release-relevant dimensions.

```{r}
#| fig-width: 14
#| fig-height: 10

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

deviation_data <- communications_integrated %>%
  mutate(
    event_text = str_to_lower(replace_na(content, "")),
    
    period = case_when(
      date < as_date("2046-06-05") ~ "Baseline before June 5",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5–6PM breach window",
      TRUE ~ NA_character_
    ),
    
    decision_control_signal = str_detect(
      event_text,
      "approve|approval|greenlight|go ahead|clearance|execute|publish|post|release|hold|wait|pause|review|monitor|block|do not|don't|cannot|not ready|judge|enforcement|guardrail|compliance|policy"
    )
  ) %>%
  filter(!is.na(period))

agent_deviation_profile <- deviation_data %>%
  group_by(agent_label, period) %>%
  summarise(
    public_share = mean(is_public, na.rm = TRUE),
    merger_share = mean(contains_merger, na.rm = TRUE),
    public_merger_share = mean(is_public & contains_merger, na.rm = TRUE),
    decision_control_share = mean(decision_control_signal, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  tidyr::complete(
    agent_label,
    period = c("Baseline before June 5", "5–6PM breach window"),
    fill = list(
      public_share = 0,
      merger_share = 0,
      public_merger_share = 0,
      decision_control_share = 0
    )
  ) %>%
  pivot_longer(
    cols = c(
      public_share,
      merger_share,
      public_merger_share,
      decision_control_share
    ),
    names_to = "metric",
    values_to = "share"
  ) %>%
  pivot_wider(
    names_from = period,
    values_from = share
  ) %>%
  mutate(
    deviation = abs(`5–6PM breach window` - `Baseline before June 5`),
    
    metric = recode(
      metric,
      public_share = "Public\nCommunication",
      merger_share = "Merger\nContent",
      public_merger_share = "Public +\nMerger",
      decision_control_share = "Decision /\nControl"
    ),
    
    metric = factor(
      metric,
      levels = c(
        "Public\nCommunication",
        "Merger\nContent",
        "Public +\nMerger",
        "Decision /\nControl"
      )
    )
  )

agent_deviation_score <- agent_deviation_profile %>%
  group_by(agent_label) %>%
  summarise(
    deviation_score = sum(deviation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(deviation_score))

agent_deviation_profile <- agent_deviation_profile %>%
  left_join(agent_deviation_score, by = "agent_label") %>%
  mutate(
    agent_label = factor(
      agent_label,
      levels = rev(agent_deviation_score$agent_label)
    )
  )

# Scale all deviations to the same 0–1 range for radar plotting
max_deviation <- max(agent_deviation_profile$deviation, na.rm = TRUE)

radar_data <- agent_deviation_profile %>%
  mutate(
    scaled_deviation = if (is.na(max_deviation) | max_deviation == 0) {
      0
    } else {
      deviation / max_deviation
    },
    
    metric_id = as.integer(metric),
    angle = 2 * pi * (metric_id - 1) / nlevels(metric),
    x = scaled_deviation * sin(angle),
    y = scaled_deviation * cos(angle),
    
    label_x = 1.18 * sin(angle),
    label_y = 1.18 * cos(angle),
    
    value_label = ifelse(
      deviation > 0,
      scales::percent(deviation, accuracy = 1),
      ""
    )
  )

# Close each radar polygon
radar_polygon_data <- radar_data %>%
  arrange(agent_label, metric_id) %>%
  group_by(agent_label) %>%
  bind_rows(slice(., 1)) %>%
  ungroup()

# Radar grid
grid_levels <- c(0.25, 0.5, 0.75, 1)

radar_grid <- expand_grid(
  grid_level = grid_levels,
  metric_id = 1:nlevels(radar_data$metric)
) %>%
  mutate(
    angle = 2 * pi * (metric_id - 1) / nlevels(radar_data$metric),
    x = grid_level * sin(angle),
    y = grid_level * cos(angle)
  ) %>%
  group_by(grid_level) %>%
  bind_rows(slice(., 1)) %>%
  ungroup()

radar_axes <- tibble(
  metric = levels(radar_data$metric),
  metric_id = 1:nlevels(radar_data$metric)
) %>%
  mutate(
    angle = 2 * pi * (metric_id - 1) / nlevels(radar_data$metric),
    x = sin(angle),
    y = cos(angle),
    label_x = 1.08 * sin(angle),
    label_y = 1.08 * cos(angle)
  )

ggplot() +
  geom_path(
    data = radar_grid,
    aes(
      x = x,
      y = y,
      group = grid_level
    ),
    colour = "grey82",
    linewidth = 0.4
  ) +
  geom_segment(
    data = radar_axes,
    aes(
      x = 0,
      y = 0,
      xend = x,
      yend = y
    ),
    colour = "grey78",
    linewidth = 0.4
  ) +
  geom_polygon(
    data = radar_polygon_data,
    aes(
      x = x,
      y = y,
      group = agent_label
    ),
    fill = "#E15759",
    alpha = 0.25,
    colour = "#E15759",
    linewidth = 0.8
  ) +
  geom_point(
    data = radar_data,
    aes(
      x = x,
      y = y
    ),
    colour = "#E15759",
    size = 2
  ) +
  geom_text(
    data = radar_data,
    aes(
      x = x,
      y = y,
      label = value_label
    ),
    size = 2.7,
    vjust = -0.7
  ) +
  geom_text(
    data = radar_axes,
    aes(
      x = label_x,
      y = label_y,
      label = metric
    ),
    size = 2.3,
    lineheight = 0.9
  ) +
  facet_wrap(
    ~ agent_label,
    ncol = 3
  ) +
  coord_equal() +
  labs(
    title = "Behavioral Deviation Profile by Agent",
    subtitle = "Radar panels show each agent's release-window deviation from baseline behavior",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    plot.margin = margin(10, 15, 10, 15)
  )
```

```{r}
agent_deviation_score %>%
  mutate(
    `Deviation Score` = round(deviation_score, 3)
  ) %>%
  select(
    Agent = agent_label,
    `Deviation Score`
  ) %>%
  knitr::kable()
```

This radar profile shows that behavioral deviation was uneven across agents. Legal-Agent and Social-Manager-Agent show the most breach-relevant shifts because their deviations include strong increases in merger content and visible public + merger exposure. Intern-Agent and Judge-Agent also show large deviations, but these are mainly driven by decision/control or public communication changes rather than direct public-merger exposure. Overall, the release-window behavior was not a uniform system-wide shift; the most important deviations were concentrated in agents connected to merger content, control activity, and public-facing exposure.

## 7.5 Public-Merger Behavior as a New Pattern

This section focuses specifically on public merger-related communication. The purpose is to test whether public merger exposure before the embargo lift was part of normal prior behavior or a new pattern that emerged on June 5.

```{r}
#| fig-width: 11
#| fig-height: 6.5

embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")

public_merger_pattern <- communications_integrated %>%
  mutate(
    behavior_period = case_when(
      date < as_date("2046-06-05") ~ "Baseline before June 5",
      date == as_date("2046-06-05") & timestamp < breach_time ~ "June 5 Before 5PM",
      date == as_date("2046-06-05") &
        timestamp >= breach_time &
        timestamp < embargo_lift_time ~ "5–6PM Breach Window",
      date == as_date("2046-06-05") &
        timestamp >= embargo_lift_time ~ "After 6PM Embargo Lift",
      TRUE ~ NA_character_
    ),
    
    public_merger_event = is_public & contains_merger
  ) %>%
  filter(!is.na(behavior_period)) %>%
  mutate(
    behavior_period = factor(
      behavior_period,
      levels = c(
        "Baseline before June 5",
        "June 5 Before 5PM",
        "5–6PM Breach Window",
        "After 6PM Embargo Lift"
      )
    )
  )

public_merger_summary <- public_merger_pattern %>%
  group_by(behavior_period, agent_label) %>%
  summarise(
    public_merger_count = sum(public_merger_event, na.rm = TRUE),
    total_messages = n(),
    public_merger_share = public_merger_count / total_messages,
    .groups = "drop"
  )

agent_order_public_merger <- public_merger_summary %>%
  group_by(agent_label) %>%
  summarise(
    total_public_merger = sum(public_merger_count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(total_public_merger) %>%
  pull(agent_label)

public_merger_summary <- public_merger_summary %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order_public_merger),
    label = ifelse(
      public_merger_count > 0,
      paste0(public_merger_count, " posts"),
      ""
    )
  )

ggplot(
  public_merger_summary,
  aes(
    x = behavior_period,
    y = public_merger_count,
    fill = behavior_period
  )
) +
  geom_col(
    width = 0.68,
    alpha = 0.88
  ) +
  geom_text(
    aes(label = label),
    vjust = -0.35,
    size = 3.2
  ) +
  facet_wrap(
    ~ agent_label,
    ncol = 4
  ) +
  scale_fill_manual(
    values = c(
      "Baseline before June 5" = "#4E79A7",
      "June 5 Before 5PM" = "#F28E2B",
      "5–6PM Breach Window" = "#E15759",
      "After 6PM Embargo Lift" = "#59A14F"
    )
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.22))
  ) +
  labs(
    title = "Public-Merger Behavior as a New Pattern",
    subtitle = "Public merger-related posts were absent in the baseline but appeared before the embargo lift on June 5",
    x = "Behavior Period",
    y = "Number of Public + Merger Posts",
    fill = "Behavior Period"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This chart isolates the key new behavior: public merger-related communication. Before June 5, public + merger posts were absent from the baseline. On June 5, they appeared before the official 6:00 PM embargo lift, including during the 5–6PM breach window. This confirms that the inappropriate release was not a continuation of normal prior behavior, but a new public-facing merger pattern that emerged on the breach day.

## 7.6 Key Findings from Typical vs Release Behavior Comparison

- **Pre-June 5 behavior had a clear normal structure.**\
  Before the breach day, most core agents mainly communicated through internal or private channels, while public-facing activity was concentrated among PR-Intern-Agent and Intern-Agent. Public + merger behavior was absent from the baseline, which establishes an important reference point for identifying abnormal release-window behavior.

- **The release-window behavior was not a normal continuation of the baseline.**\
  During the 5–6PM breach window, several agents shifted away from their prior behavior, especially in merger-related content, public communication, public + merger exposure, and decision/control language. This indicates that the organisation had moved from routine communication into a crisis disclosure and containment mode.

- **Public + merger exposure was the most important new pattern.**\
  The baseline shows no public + merger behavior before June 5. However, on June 5, public merger-related posts appeared before the official 6PM embargo lift, especially involving Legal-Agent and Social-Manager-Agent. This confirms that public-facing merger communication was a new and abnormal pattern rather than part of normal prior behaviour.

- **The strongest breach-relevant deviations were concentrated in specific agents.**\
  Legal-Agent and Social-Manager-Agent show the most important release-window shifts because they combine merger-related content, public + merger exposure, and control-oriented communication. Intern-Agent, PR-Intern-Agent, and Judge-Agent also show deviations, but their deviations are more related to public communication or decision/control behavior rather than direct evidence of the original leakage source.

- **Part 7 shows post-leak behavioral change, not the original leak mechanism.**\
  The behavior comparison confirms that the 5–6PM release window was abnormal, but it does not by itself identify who first leaked the merger information. The Slack screenshot evidence from Part 4 provides the direct internal-leak evidence, while Part 7 shows how agents behaved after the information environment had already shifted toward public exposure and crisis management.

- **Overall, the evidence supports abnormal release-window behavior consistent with internal leak containment and post-leak crisis management.**\
  The release-window pattern differs sharply from prior behavior: merger-related content became more prominent, public + merger exposure emerged for the first time, and control-related activity remained active. Therefore, Part 7 should not be interpreted as proving a senior-orchestrated leak. Instead, it supports the broader conclusion that an internal confidentiality leak was followed by abnormal, high-pressure system behavior as agents attempted to contain, manage, and formalise the disclosure before and around the embargo lift.

# 8. Leading Indicators and System Pressure

The purpose of this section is to examine whether there were warning signs before the inappropriate release occurred. While Part 7 showed that the release-window behavior differed from normal prior behavior, this section focuses on whether the system was already showing signs of pressure, sensitivity, or instability before the embargo boundary was crossed.

This analysis focuses on the first part of Task 3: whether there were leading indicators that such a release was possible. Leading indicators are defined as communication patterns that appeared before the official embargo lift and suggested elevated release risk, including urgency signals, merger-related content, confidentiality references, risk language, error or failure language, and public-facing exposure.

The section does not yet analyse prior anomalies or explain why earlier warning signs did not trigger intervention. Those questions are addressed separately in Part 9. Instead, Part 8 focuses on identifying the pressure environment around the breach and assessing whether the June 5 release was preceded by measurable warning signals.

## 8.1 System Pressure Indicators

```{r}
pressure_data <- communications_integrated %>%
  mutate(
    risk_signal = str_detect(
      content_lower,
      "risk|exposure|liability|regulator|regulatory|investigation|scrutiny|concern|threat"
    ),
    
    urgency_signal = str_detect(
      content_lower,
      "urgent|asap|immediately|deadline|pressure|crisis|escalate|fast|now|critical"
    ),
    
    error_signal = str_detect(
      content_lower,
      "error|failed|failure|mistake|confusion|unclear|misfire|wrong|breakdown|problem"
    ),
    
    confidentiality_signal = str_detect(
      content_lower,
      "confidential|embargo|private|internal only|not public|restricted|sensitive"
    ),
    
    merger_signal = contains_merger,
    
    leak_exposure_signal = str_detect(
      content_lower,
      "leak|leaked|leaking|screenshot|slack|said-it|saidit|reddit|subreddit|public exposure|information perimeter|perimeter|breach|breached|unauthorized|unauthorised|disclosure|exposed|surfaced"
    ),
    
    any_pressure_signal = risk_signal |
      urgency_signal |
      error_signal |
      confidentiality_signal |
      merger_signal |
      leak_exposure_signal,
    
    pressure_score = 
      as.integer(risk_signal) +
      as.integer(urgency_signal) +
      as.integer(error_signal) +
      as.integer(confidentiality_signal) +
      as.integer(merger_signal) +
      as.integer(leak_exposure_signal)
  )
```

```{r}
pressure_indicator_summary <- tibble(
  Pressure_Indicator = c(
    "Risk Signal",
    "Urgency Signal",
    "Error / Failure Signal",
    "Confidentiality Signal",
    "Merger Signal",
    "Leak / Exposure Signal",
    "Any Pressure Signal"
  ),
  Definition = c(
    "Messages containing risk, exposure, liability, regulatory, investigation, scrutiny, concern, or threat-related language",
    "Messages containing urgent, deadline, crisis, escalation, or immediate-action language",
    "Messages containing error, failure, confusion, mistake, or breakdown-related language",
    "Messages containing confidential, embargo, private, restricted, or sensitive-information language",
    "Messages containing merger, acquisition, CivicLoom, HarborCrest, deal, or transaction-related language",
    "Messages containing leak, screenshot, Slack, Said-it, public exposure, information perimeter, breach, or unauthorised-disclosure language",
    "Messages containing at least one of the above pressure-related indicators"
  ),
  Message_Count = c(
    sum(pressure_data$risk_signal, na.rm = TRUE),
    sum(pressure_data$urgency_signal, na.rm = TRUE),
    sum(pressure_data$error_signal, na.rm = TRUE),
    sum(pressure_data$confidentiality_signal, na.rm = TRUE),
    sum(pressure_data$merger_signal, na.rm = TRUE),
    sum(pressure_data$leak_exposure_signal, na.rm = TRUE),
    sum(pressure_data$any_pressure_signal, na.rm = TRUE)
  )
)

pressure_indicator_summary
```

The pressure indicator summary shows that the crisis environment was dominated by **urgency**, **merger-related communication**, and **risk-related language**. Among all categories, Urgency Signal has the highest count with 413 messages, suggesting that agents were operating under strong time pressure and immediate-response demands. Merger Signal is also highly visible with 243 messages, confirming that merger-related information was widely circulating during the crisis period. Importantly, the newly added Leak / Exposure Signal appears in 120 messages, showing that leakage-related language was not marginal; it became a meaningful part of the system pressure environment. 

Overall, the table supports the interpretation that the breach did not occur in a normal communication setting. Instead, it happened under concentrated urgency, merger sensitivity, confidentiality pressure, and direct leak/exposure concerns. This strengthens the mixed conclusion that the Slack-related internal leak became a wider disclosure crisis because the multi-agent communication system failed to contain sensitive information under pressure.

## 8.2 Pressure Signal Heatmap Over Time

This visual examines how pressure signals were distributed across time. Instead of using a simple line chart, the heatmap highlights which pressure categories became more concentrated on specific dates.

```{r}
pressure_heatmap_data <- pressure_data %>%
  group_by(date) %>%
  summarise(
    Risk = sum(risk_signal, na.rm = TRUE),
    Urgency = sum(urgency_signal, na.rm = TRUE),
    `Error / Failure` = sum(error_signal, na.rm = TRUE),
    Confidentiality = sum(confidentiality_signal, na.rm = TRUE),
    Merger = sum(merger_signal, na.rm = TRUE),
    `Leak / Exposure` = sum(leak_exposure_signal, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(Risk, Urgency, `Error / Failure`, Confidentiality, Merger, `Leak / Exposure`),
    names_to = "pressure_signal",
    values_to = "message_count"
  )

ggplot(
  pressure_heatmap_data,
  aes(
    x = date,
    y = pressure_signal,
    fill = message_count
  )
) +
  geom_tile(
    color = "white",
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = ifelse(message_count > 0, message_count, "")),
    size = 3
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkred"
  ) +
  labs(
    title = "Pressure Signal Heatmap Over Time",
    subtitle = "Darker cells indicate stronger concentration of pressure-related communications",
    x = "Date",
    y = "Pressure Signal",
    fill = "Message Count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

The pressure signal heatmap shows that June 5 was the decisive escalation point in the crisis. Before June 5, pressure-related communication was scattered and relatively low across urgency, risk, merger, confidentiality, error/failure, and leak/exposure signals. On June 5, all major pressure categories intensified sharply, with urgency and merger-related signals forming the strongest concentrations. The newly added Leak / Exposure row is important because it separates direct leakage language from general risk or confidentiality pressure.

Overall, this heatmap supports the interpretation that system pressure amplified the internal leak rather than replacing it as the explanation. The Slack-related exposure occurred in an environment where urgency, merger sensitivity, legal risk, confidentiality pressure, and leak/exposure concerns converged, making the breach harder to contain and pushing agents toward accelerated post-leak crisis management.

## 8.3 Pressure Signal Composition

This visual shows the composition of pressure-related signals across the communication dataset. The donut chart is used because the purpose here is to compare the relative share of each pressure source rather than track changes over time.

```{r}
pressure_composition <- pressure_data %>%
  summarise(
    Risk = sum(risk_signal, na.rm = TRUE),
    Urgency = sum(urgency_signal, na.rm = TRUE),
    `Error / Failure` = sum(error_signal, na.rm = TRUE),
    Confidentiality = sum(confidentiality_signal, na.rm = TRUE),
    Merger = sum(merger_signal, na.rm = TRUE),
    `Leak / Exposure` = sum(leak_exposure_signal, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "pressure_signal",
    values_to = "message_count"
  ) %>%
  mutate(
    share = message_count / sum(message_count),
    label = paste0(
      pressure_signal,
      "\n",
      message_count,
      " messages",
      "\n",
      scales::percent(share, accuracy = 0.1)
    )
  ) %>%
  arrange(desc(message_count)) %>%
  mutate(
    pressure_signal = factor(pressure_signal, levels = pressure_signal),
    ymax = cumsum(share),
    ymin = lag(ymax, default = 0),
    label_position = (ymax + ymin) / 2
  )

ggplot(
  pressure_composition,
  aes(
    ymax = ymax,
    ymin = ymin,
    xmax = 4,
    xmin = 2.6,
    fill = pressure_signal
  )
) +
  geom_rect(
    colour = "white",
    linewidth = 0.8
  ) +
  coord_polar(theta = "y") +
  xlim(c(1.5, 4.5)) +
  geom_text(
    aes(
      x = 4.25,
      y = label_position,
      label = label
    ),
    size = 2.8,
    lineheight = 0.9
  ) +
  labs(
    title = "Pressure Signal Composition",
    subtitle = "Share of each pressure signal category",
    fill = "Pressure Signal"
  ) +
  theme_void() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 18),
    plot.subtitle = element_text(size = 11)
  )
```

The pressure signal composition chart shows that **urgency was the dominant pressure category**, accounting for 413 messages, or 38.9% of all pressure signals. This suggests that agents were operating under strong time pressure and immediate-response demands during the crisis period. **Merger-related communication** was the second-largest category, with 243 messages, or 22.9%, confirming that merger information was highly active in the communication environment.

The newly added **Leak / Exposure** category is also important. It accounts for 120 messages, or 11.3% of all pressure signals. This shows that direct leakage-related language was not marginal. The crisis was not only about general urgency or merger discussion; it also involved explicit exposure signals linked to screenshots, public leakage, and information-boundary failure.

Overall, this graph supports the revised interpretation that **system pressure amplified an internal confidentiality leak**. Urgency and merger signals explain why the communication system was under stress, while the Leak / Exposure category captures the direct breach-related language. Therefore, the pressure pattern is consistent with a mixed pathway: **internal leak, system-level control failure, and post-leak crisis management**.

## 8.4 Agent-Level Pressure Exposure

This visual examines how pressure signals are distributed across agents. The heatmap helps identify which agents were most exposed to urgency, risk, merger, confidentiality, or error-related communication.

```{r}
agent_pressure_exposure <- pressure_data %>%
  group_by(agent_label) %>%
  summarise(
    Risk = sum(risk_signal, na.rm = TRUE),
    Urgency = sum(urgency_signal, na.rm = TRUE),
    `Error / Failure` = sum(error_signal, na.rm = TRUE),
    Confidentiality = sum(confidentiality_signal, na.rm = TRUE),
    Merger = sum(merger_signal, na.rm = TRUE),
    `Leak / Exposure` = sum(leak_exposure_signal, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(Risk, Urgency, `Error / Failure`, Confidentiality, Merger, `Leak / Exposure`),
    names_to = "pressure_signal",
    values_to = "message_count"
  )

agent_order <- agent_pressure_exposure %>%
  group_by(agent_label) %>%
  summarise(total_pressure = sum(message_count), .groups = "drop") %>%
  arrange(total_pressure) %>%
  pull(agent_label)

agent_pressure_exposure <- agent_pressure_exposure %>%
  mutate(
    agent_label = factor(agent_label, levels = agent_order),
    pressure_signal = factor(
      pressure_signal,
      levels = c("Urgency", "Merger", "Risk", "Confidentiality", "Error / Failure","Leak / Exposure")
    )
  )

ggplot(
  agent_pressure_exposure,
  aes(
    x = pressure_signal,
    y = agent_label,
    fill = message_count
  )
) +
  geom_tile(
    color = "white",
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = ifelse(message_count > 0, message_count, "")),
    size = 3.5
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkred"
  ) +
  labs(
    title = "Agent-Level Pressure Exposure",
    subtitle = "Darker cells indicate higher exposure to pressure-related communication",
    x = "Pressure Signal",
    y = "Agent",
    fill = "Message Count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 25, hjust = 1)
  )
```

The agent-level pressure exposure heatmap shows that crisis pressure was concentrated around a small group of core agents rather than distributed evenly across the whole system. Legal-Agent and Social-Manager-Agent had the highest exposure to urgency and merger-related pressure, indicating that they were most directly involved in interpreting, monitoring, and responding to the breach as it escalated. Legal-Agent also had the highest exposure to risk, confidentiality, and error/failure signals, which confirms its central role in assessing legal exposure and information-control failure.

The newly added Leak / Exposure column is especially important. Social-Manager-Agent and Legal-Agent show the highest leak/exposure counts, with 54 and 40 messages respectively. This suggests that the agents closest to public monitoring and legal interpretation were also the agents most exposed to direct leakage concerns. Platform-Trust-Agent had lower leak/exposure exposure but remained important through urgency and risk signals, reflecting its coordination and system-control role.

Overall, this heatmap supports the revised interpretation that the breach was not simply a general system breakdown. The pressure was concentrated around agents responsible for legal escalation, public monitoring, and crisis coordination. This means the Slack-related internal leak entered a communication environment where the key control agents were already under intense urgency, merger, risk, and confidentiality pressure, making containment more difficult and pushing the system toward accelerated post-leak crisis management.

## 8.5 Leading Indicator Score Before the Breach

This section examines whether there were warning signs before the embargo breach. Instead of looking at pressure signals separately, a combined leading indicator score is constructed from several risk-relevant conditions: merger-related content, confidentiality signals, urgency signals, public-facing activity, and boundary-agent involvement.

```{r}
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")

# Identify boundary agents based on communication scope
boundary_agent_profile <- pressure_data %>%
  count(agent_label, communication_scope, name = "message_count") %>%
  group_by(agent_label) %>%
  mutate(
    total_messages = sum(message_count),
    share = message_count / total_messages
  ) %>%
  ungroup() %>%
  select(agent_label, communication_scope, share) %>%
  pivot_wider(
    names_from = communication_scope,
    values_from = share,
    values_fill = 0
  )

if (!"Public" %in% names(boundary_agent_profile)) {
  boundary_agent_profile$Public <- 0
}

if (!"Private / One-on-One" %in% names(boundary_agent_profile)) {
  boundary_agent_profile[["Private / One-on-One"]] <- 0
}

boundary_agents <- boundary_agent_profile %>%
  mutate(
    boundary_bridge_score = Public + `Private / One-on-One`
  ) %>%
  filter(
    boundary_bridge_score >= median(boundary_bridge_score, na.rm = TRUE)
  ) %>%
  pull(agent_label)

# Build message-level leading indicator components
leading_indicator_messages <- pressure_data %>%
  filter(timestamp < embargo_lift_time) %>%
  mutate(
    public_merger_signal = is_public & merger_signal,
    
    public_leak_exposure_signal = is_public & leak_exposure_signal,
    
    merger_confidential_signal = merger_signal & confidentiality_signal,
    
    merger_urgency_signal = merger_signal & urgency_signal,
    
    merger_leak_exposure_signal = merger_signal & leak_exposure_signal,
    
    public_pressure_signal = is_public & any_pressure_signal,
    
    boundary_agent_pressure_signal = agent_label %in% boundary_agents &
      (
        merger_signal |
          confidentiality_signal |
          urgency_signal |
          leak_exposure_signal
      ),
    
    leading_indicator_score =
      3 * as.integer(public_merger_signal) +
      3 * as.integer(public_leak_exposure_signal) +
      2 * as.integer(merger_confidential_signal) +
      2 * as.integer(merger_urgency_signal) +
      2 * as.integer(merger_leak_exposure_signal) +
      2 * as.integer(public_pressure_signal) +
      1 * as.integer(boundary_agent_pressure_signal)
  )

# Aggregate by date
daily_leading_score <- leading_indicator_messages %>%
  group_by(date) %>%
  summarise(
    leading_indicator_score = sum(leading_indicator_score, na.rm = TRUE),
    public_merger_messages = sum(public_merger_signal, na.rm = TRUE),
    public_leak_exposure_messages = sum(public_leak_exposure_signal, na.rm = TRUE),
    merger_confidential_messages = sum(merger_confidential_signal, na.rm = TRUE),
    merger_urgency_messages = sum(merger_urgency_signal, na.rm = TRUE),
    merger_leak_exposure_messages = sum(merger_leak_exposure_signal, na.rm = TRUE),
    public_pressure_messages = sum(public_pressure_signal, na.rm = TRUE),
    boundary_agent_pressure_messages = sum(boundary_agent_pressure_signal, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    public_breach_related_messages =
      public_merger_messages + public_leak_exposure_messages
  )

max_score <- max(daily_leading_score$leading_indicator_score, na.rm = TRUE)

daily_leading_score <- daily_leading_score %>%
  mutate(
    relative_score = if (max_score == 0) {
      0
    } else {
      leading_indicator_score / max_score * 100
    },
    
    warning_level = case_when(
      relative_score >= 75 ~ "Critical",
      relative_score >= 40 ~ "High",
      relative_score >= 10 ~ "Moderate",
      TRUE ~ "Low"
    ),
    
    warning_level = factor(
      warning_level,
      levels = c("Low", "Moderate", "High", "Critical")
    )
  )

ggplot(
  daily_leading_score,
  aes(
    x = date,
    y = relative_score
  )
) +
  geom_segment(
    aes(
      xend = date,
      y = 0,
      yend = relative_score
    ),
    colour = "grey70",
    linewidth = 0.7
  ) +
  geom_point(
    aes(
      size = public_breach_related_messages + 1,
      fill = warning_level
    ),
    shape = 21,
    colour = "grey30",
    alpha = 0.85
  ) +
  geom_text(
    aes(
      label = ifelse(leading_indicator_score > 0, leading_indicator_score, "")
    ),
    vjust = -1,
    size = 3
  ) +
  scale_y_continuous(
    limits = c(0, 110),
    breaks = seq(0, 100, 25),
    labels = function(x) paste0(x, "%")
  ) +
  scale_size_continuous(
    range = c(3, 12),
    name = "Public Merger or\nLeak Exposure\nMessages + 1"
  ) +
  labs(
    title = "Leading Indicator Score Before the Breach",
    subtitle = "Relative warning score based on merger, confidentiality, urgency, leak/exposure, public exposure, and boundary-agent activity",
    x = "Date",
    y = "Relative Leading Indicator Score",
    fill = "Warning Level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

The updated leading indicator chart incorporates leak/exposure signals into the warning score. This makes the indicator more aligned with the revised interpretation of the incident, because the breach was not only associated with merger, confidentiality, urgency, and public exposure signals, but also with direct leakage-related language such as screenshots, Slack, Said-it, and compromised information perimeter references.

The chart shows that June 5 remained the decisive warning outlier before the embargo lift. Earlier dates contained only scattered and low-level warning signals, while June 5 shows a sharp spike in the relative leading indicator score. This supports a missed-intervention interpretation: warning conditions were visible before the final breach window, but they were not converted into effective containment action. The chart does not prove a senior-orchestrated leak; instead, it shows that an internal leak became consequential because the wider communication system failed to respond to escalating warning indicators in time.

```{r}
leading_indicator_component_table <- daily_leading_score %>%
  arrange(desc(leading_indicator_score)) %>%
  transmute(
    Date = date,
    `Leading Indicator Score` = leading_indicator_score,
    `Relative Score` = paste0(round(relative_score, 1), "%"),
    `Public Merger Messages` = public_merger_messages,
    `Public Leak / Exposure Messages` = public_leak_exposure_messages,
    `Public Merger or Leak / Exposure Messages` = public_breach_related_messages,
    `Merger + Confidentiality` = merger_confidential_messages,
    `Merger + Urgency` = merger_urgency_messages,
    `Merger + Leak / Exposure` = merger_leak_exposure_messages,
    `Public Pressure Messages` = public_pressure_messages,
    `Boundary-Agent Pressure Messages` = boundary_agent_pressure_messages,
    `Warning Level` = warning_level
  )

DT::datatable(
  leading_indicator_component_table,
  rownames = FALSE,
  options = list(
    pageLength = 8,
    scrollX = TRUE
  )
)
```

## 8.7 Key Findings from System Pressure Analysis

- **System pressure was substantial and concentrated.**\
  A total of 555 messages contained at least one pressure signal. Urgency was the dominant pressure category, with 413 messages, followed by merger-related signals with 243 messages and risk signals with 150 messages. This shows that the breach occurred in a highly pressured communication environment rather than during a stable operating period.

- **June 5 was the clear pressure escalation point.**\
  The pressure heatmap shows that urgency, merger, risk, confidentiality, and error/failure signals all intensified sharply on June 5. This indicates that the breach day was a system-wide escalation point, where sensitive merger information, legal exposure, public attention, and communication urgency converged at the same time.

- **Pressure was mainly driven by urgency and merger escalation.**\
  The composition chart shows that urgency accounted for 43.8% of all pressure signals, while merger-related signals accounted for 25.8%. Error/failure language was much smaller, suggesting that the problem was less about a technical malfunction and more about urgency-driven governance pressure, confidentiality stress, and failure to contain sensitive information once the leak risk had materialised.

- **Pressure exposure was concentrated around core agents.**\
  Legal-Agent and Social-Manager-Agent had the highest exposure to urgency and merger-related pressure. Platform-Trust-Agent also showed strong risk and urgency exposure, making it important as a coordination and system-control actor. This pattern suggests that the agents closest to legal interpretation, public monitoring, and crisis coordination were also operating under the strongest pressure.

- **Leading indicators existed before the final breach, but June 5 was the decisive outlier.**\
  Earlier dates showed small warning scores, while June 5 jumped to a leading indicator score of 711. This suggests that weak warning signals existed before the breach, but multiple risk conditions converged sharply on the breach day. These indicators help explain why the internal Slack leak and public merger exposure became difficult to contain once the crisis accelerated.

- **System pressure should be interpreted as an enabling condition, not the sole cause of the breach.**\
  The Part 8 visuals show escalating urgency, repeated warning patterns, and concentrated pressure around sensitive merger communication. However, this does not mean the breach was purely a system breakdown. Instead, the pressure environment made the internal confidentiality leak more damaging by weakening containment, accelerating public-facing responses, and forcing agents into rapid disclosure-management decisions.

- **Overall, Part 8 supports the broader mixed interpretation: internal leak plus system-level control failure.**\
  The evidence does not point to a normal communication environment or a single routine posting error. It shows a communication system under concentrated urgency, merger sensitivity, confidentiality pressure, and public-facing exposure. Therefore, Part 8 supports the conclusion that the Slack-related internal leak became a wider disclosure crisis because the multi-agent system failed to contain it under pressure, followed by deliberate post-leak crisis management.

# 9. Prior Anomalies and Missed Intervention

## 9.1 Purpose of Prior Anomaly Analysis

The purpose of this section is to examine whether the system showed earlier warning patterns before the final embargo breach, and why those earlier signals did not result in visible intervention. While Part 8 focuses on leading indicators and system pressure around the breach, Part 9 looks backward to identify prior anomalies, near-miss behaviors, and missed opportunities for control.

This section addresses the remaining parts of Task 3:

1.  Were there prior occasions where agents’ actual behaviors differed from expected behavior?
2.  Were there other occasions where the agent system exhibited behaviors similar to those that led to the inappropriate release?
3.  Why did those prior occasions not result in noticeable action?

The analysis uses the typical behavior baseline from Part 7 as the reference point for “expected behavior.” Prior anomalies are defined as earlier communication records or patterns that showed partial similarity to the final breach pathway, such as merger-related pressure, confidentiality references, public-facing pressure, decision/control activity, or near-public exposure.

The goal is not to claim that earlier anomalies directly caused the breach. Instead, this section evaluates whether the system had warning signs that could have been detected earlier, and whether those signs were too weak, too scattered, or too poorly escalated to trigger effective intervention before the June 5 release window.

## 9.2 Prior Warning Occasions and Near-Miss Signals

```{r}
embargo_lift_time <- ymd_hms("2046-06-05 18:00:00")
breach_time <- ymd_hms("2046-06-05 17:00:00")
breach_date <- as.Date("2046-06-05")

near_miss_data <- pressure_data %>%
  filter(timestamp < embargo_lift_time) %>%
  mutate(
    date_fixed = as.Date(timestamp),
    
    pre_breach_phase = case_when(
      date_fixed < breach_date ~ "Prior to June 5",
      timestamp < breach_time ~ "June 5 Before 5:00 PM",
      timestamp >= breach_time & timestamp < embargo_lift_time ~ "5:00–6:00 PM Breach Window",
      TRUE ~ NA_character_
    ),
    
    near_miss_pattern = case_when(
      # Most critical: public-facing breach-related signals
      is_public & merger_signal ~ "Public + Merger",
      is_public & leak_exposure_signal ~ "Public + Leak / Exposure",
      
      # Direct leakage combined with merger/confidentiality/urgency
      merger_signal & leak_exposure_signal ~ "Merger + Leak / Exposure",
      confidentiality_signal & leak_exposure_signal ~ "Leak / Exposure + Confidentiality",
      urgency_signal & leak_exposure_signal ~ "Leak / Exposure + Urgency",
      
      # Existing near-miss patterns
      is_public & any_pressure_signal ~ "Public + Pressure",
      merger_signal & confidentiality_signal ~ "Merger + Confidentiality",
      merger_signal & urgency_signal ~ "Merger + Urgency",
      merger_signal & risk_signal ~ "Merger + Risk",
      confidentiality_signal & urgency_signal ~ "Confidentiality + Urgency",
      
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(pre_breach_phase), !is.na(near_miss_pattern)) %>%
  mutate(
    near_miss_severity = case_when(
      near_miss_pattern %in% c(
        "Public + Merger",
        "Public + Leak / Exposure",
        "Merger + Leak / Exposure"
      ) ~ "Critical",
      
      near_miss_pattern %in% c(
        "Public + Pressure",
        "Merger + Confidentiality",
        "Merger + Urgency",
        "Leak / Exposure + Confidentiality",
        "Leak / Exposure + Urgency"
      ) ~ "High",
      
      TRUE ~ "Moderate"
    ),
    
    near_miss_severity = factor(
      near_miss_severity,
      levels = c("Moderate", "High", "Critical")
    ),
    
    pre_breach_phase = factor(
      pre_breach_phase,
      levels = c(
        "Prior to June 5",
        "June 5 Before 5:00 PM",
        "5:00–6:00 PM Breach Window"
      )
    )
  )
```

```{r}
near_miss_summary <- near_miss_data %>%
  count(
    pre_breach_phase,
    near_miss_pattern,
    near_miss_severity,
    name = "event_count"
  ) %>%
  mutate(
    pre_breach_phase = factor(
      pre_breach_phase,
      levels = c(
        "Prior to June 5",
        "June 5 Before 5:00 PM",
        "5:00–6:00 PM Breach Window"
      )
    )
  )

ggplot(
  near_miss_summary,
  aes(
    x = pre_breach_phase,
    y = near_miss_pattern,
    size = event_count,
    fill = near_miss_severity
  )
) +
  geom_point(
    shape = 21,
    colour = "grey30",
    alpha = 0.85
  ) +
  geom_text(
    aes(label = event_count),
    size = 3,
    colour = "black"
  ) +
  scale_size_continuous(
    range = c(4, 16),
    name = "Event Count"
  ) +
  labs(
    title = "Prior Warning Occasions and Near-Miss Signals",
    subtitle = "Pre-embargo risk patterns by phase and severity",
    x = "Pre-Embargo Phase",
    y = "Near-Miss Pattern",
    fill = "Severity"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 20, hjust = 1)
  )
```

The updated near-miss chart shows that warning signals were already visible before the official embargo lift, but they intensified sharply on June 5. The largest concentration appeared in the June 5 before-5PM phase, especially around `Merger + Leak / Exposure`, `Merger + Urgency`, and `Leak / Exposure + Urgency`. This indicates that the system was already showing signs of direct exposure risk before the 5:00–6:00 PM breach window.

The chart also shows that the breach window still contained critical public-facing patterns, including `Public + Merger` and `Merger + Leak / Exposure`. This suggests that the final breach was not a sudden isolated posting mistake. Instead, it followed a sequence of near-miss signals in which merger sensitivity, urgency, confidentiality pressure, and leak/exposure language accumulated before the embargo lift.

Overall, this chart supports a missed-intervention interpretation. Earlier warning patterns existed, but they were not converted into a strong enough containment response. The Slack-related internal leak became consequential because leak/exposure signals and public-facing merger risk were allowed to escalate into the breach window.

## 9.3 Near-Miss Signals by Agent

```{r}
near_miss_event_table <- near_miss_data %>%
  arrange(desc(near_miss_severity), timestamp) %>%
  transmute(
    Time = timestamp,
    Phase = pre_breach_phase,
    Severity = near_miss_severity,
    Pattern = near_miss_pattern,
    Agent = agent_label,
    Role = agent_role,
    Channel = channel,
    Scope = communication_scope,
    Message_Type = message_type,
    Content = str_trunc(content, width = 220)
  )

DT::datatable(
  near_miss_event_table,
  rownames = FALSE,
  options = list(
    pageLength = 8,
    lengthMenu = c(8, 15, 30),
    scrollX = TRUE,
    scrollY = "360px",
    autoWidth = FALSE,
    columnDefs = list(
      list(width = "130px", targets = 0),
      list(width = "170px", targets = 1),
      list(width = "100px", targets = 2),
      list(width = "180px", targets = 3),
      list(width = "150px", targets = 4),
      list(width = "110px", targets = 5),
      list(width = "130px", targets = 6),
      list(width = "140px", targets = 7),
      list(width = "130px", targets = 8),
      list(width = "560px", targets = 9)
    )
  )
) %>%
  DT::formatStyle(
    "Content",
    `white-space` = "normal",
    `font-size` = "12px",
    `line-height` = "1.25"
  )
```

```{r}
near_miss_agent_summary <- near_miss_data %>%
  count(
    agent_label,
    near_miss_pattern,
    name = "event_count"
  ) %>%
  group_by(agent_label) %>%
  mutate(
    total_near_miss_events = sum(event_count)
  ) %>%
  ungroup()

ggplot(
  near_miss_agent_summary,
  aes(
    x = reorder(agent_label, total_near_miss_events),
    y = event_count,
    fill = near_miss_pattern
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Near-Miss Signals by Agent",
    subtitle = "Agents involved in pre-embargo warning patterns",
    x = "Agent",
    y = "Number of Near-Miss Signals",
    fill = "Near-Miss Pattern"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
```

This chart shows that near-miss warning signals were concentrated mainly around **Legal-Agent** and **Social-Manager-Agent**. Legal-Agent had the highest overall exposure, especially across merger-related, leak/exposure, and confidentiality-linked patterns, which confirms its central role in legal escalation and breach assessment. Social-Manager-Agent also carried substantial near-miss signals, reflecting its role in monitoring public exposure and detecting external discussion. Overall, the warning pattern was not evenly distributed across all agents; it was concentrated around the agents closest to legal control and public-facing crisis monitoring.

## 9.4 Expected vs Actual Behavior Deviations

This section compares the system’s expected behavior with the actual behaviors observed during prior warning and near-miss occasions. Expected behavior is based on the pre-June 5 baseline established in Part 7.

```{r}
#| fig-width: 10.5
#| fig-height: 6.3

expected_actual_score <- tibble(
  expected_behavior = c(
    "No public merger content",
    "Merger content contained internally",
    "Leak / exposure signals remain contained",
    "Confidentiality triggers containment",
    "Control signals reduce exposure",
    "Public-facing agents avoid sensitive amplification"
  ),
  
  `Prior to June 5` = c(
    "Low",
    "Moderate",
    "Low",
    "Moderate",
    "Moderate",
    "Moderate"
  ),
  
  `June 5 Before 5PM` = c(
    "Critical",
    "High",
    "Critical",
    "High",
    "High",
    "High"
  ),
  
  `5–6PM Breach Window` = c(
    "Critical",
    "High",
    "Critical",
    "High",
    "Critical",
    "Critical"
  )
)

expected_actual_long <- expected_actual_score %>%
  pivot_longer(
    cols = c(
      `Prior to June 5`,
      `June 5 Before 5PM`,
      `5–6PM Breach Window`
    ),
    names_to = "actual_phase",
    values_to = "deviation_level"
  ) %>%
  mutate(
    expected_behavior = factor(
      expected_behavior,
      levels = rev(expected_actual_score$expected_behavior)
    ),
    
    actual_phase = factor(
      actual_phase,
      levels = c(
        "Prior to June 5",
        "June 5 Before 5PM",
        "5–6PM Breach Window"
      )
    ),
    
    deviation_level = factor(
      deviation_level,
      levels = c(
        "Low",
        "Moderate",
        "High",
        "Critical"
      )
    ),
    
    deviation_score = case_when(
      deviation_level == "Low" ~ 1,
      deviation_level == "Moderate" ~ 2,
      deviation_level == "High" ~ 3,
      deviation_level == "Critical" ~ 4
    )
  )

ggplot(
  expected_actual_long,
  aes(
    x = actual_phase,
    y = expected_behavior,
    fill = deviation_level
  )
) +
  geom_tile(
    colour = "white",
    linewidth = 1
  ) +
  geom_text(
    aes(label = deviation_level),
    size = 4,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Low" = "#f2f2f2",
      "Moderate" = "#fdae6b",
      "High" = "#f1695b",
      "Critical" = "#a50f15"
    ),
    drop = FALSE
  ) +
  labs(
    title = "Expected vs Actual Behavior Deviations",
    subtitle = "Deviation severity increased as leak/exposure and public merger risks approached the breach window",
    x = "Actual Observation Phase",
    y = "Expected Behavior",
    fill = "Deviation Level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This deviation heatmap shows that actual behavior moved further away from expected containment as the system approached the breach window. The most serious deviations appeared on June 5, when public merger content became visible and leak/exposure signals were no longer contained. This is important because the breach was not only a failure to keep merger discussion internal; it also involved direct exposure of internal information through Slack-related screenshots and public discussion. Overall, the chart supports a missed-intervention interpretation: expected containment mechanisms existed, but they did not prevent leak/exposure signals from escalating before the embargo lift.

## 9.5 Why Prior Warnings Did Not Trigger Action

This section explains why earlier warning signals did not lead to visible intervention before the final embargo breach. The goal is to identify the main barriers that prevented prior warning patterns from becoming clear control actions.

```{r}
#| fig-width: 10.5
#| fig-height: 6.2

missed_intervention_matrix <- tibble(
  failure_reason = c(
    "Signals were scattered",
    "Risk looked contained",
    "Warnings were ambiguous",
    "Controls lacked stopping power",
    "Responsibility was distributed",
    "Escalation became clear late"
  ),
  
  `Low Signal Clarity` = c(
    "High", "Moderate", "High", "Moderate", "Moderate", "Low"
  ),
  
  `Weak Action Trigger` = c(
    "High", "High", "High", "High", "Moderate", "Moderate"
  ),
  
  `Distributed Ownership` = c(
    "Moderate", "Low", "Moderate", "High", "High", "Moderate"
  ),
  
  `Late Visibility` = c(
    "Moderate", "High", "Moderate", "Moderate", "Moderate", "High"
  )
)

missed_intervention_long <- missed_intervention_matrix %>%
  pivot_longer(
    cols = c(
      `Low Signal Clarity`,
      `Weak Action Trigger`,
      `Distributed Ownership`,
      `Late Visibility`
    ),
    names_to = "intervention_barrier",
    values_to = "barrier_level"
  ) %>%
  mutate(
    failure_reason = factor(
      failure_reason,
      levels = rev(missed_intervention_matrix$failure_reason)
    ),
    
    intervention_barrier = factor(
      intervention_barrier,
      levels = c(
        "Low Signal Clarity",
        "Weak Action Trigger",
        "Distributed Ownership",
        "Late Visibility"
      )
    ),
    
    barrier_level = factor(
      barrier_level,
      levels = c("Low", "Moderate", "High")
    )
  )

ggplot(
  missed_intervention_long,
  aes(
    x = intervention_barrier,
    y = failure_reason,
    fill = barrier_level
  )
) +
  geom_tile(
    colour = "white",
    linewidth = 1
  ) +
  geom_text(
    aes(label = barrier_level),
    size = 4,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Low" = "#f2f2f2",
      "Moderate" = "#fdae6b",
      "High" = "#a50f15"
    )
  ) +
  labs(
    title = "Why Prior Warnings Did Not Trigger Action",
    subtitle = "Intervention barriers that prevented earlier warning signs from becoming effective control actions",
    x = "Intervention Barrier",
    y = "Reason Warning Did Not Trigger Action",
    fill = "Barrier Level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold", angle = 20, hjust = 1),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11)
  )
```

This chart explains why earlier warnings did not translate into intervention. The strongest barriers were weak action triggers and low signal clarity: the warnings existed, but they were scattered and ambiguous rather than forming one clear stop signal. Risk also appeared contained before June 5 because public merger exposure was not yet obvious. By the time visibility increased, the system was already close to the breach window, so intervention came too late to prevent public exposure.

## 9.6 Key Findings from Prior Anomalies and Missed Intervention

- **Prior warning signals existed before the final breach window.**  
  The near-miss analysis shows that warning patterns were already visible before 5:00 PM on June 5. Earlier signals included `Merger + Urgency`, `Merger + Confidentiality`, and `Confidentiality + Urgency`, but the revised analysis also shows leak-related warning patterns such as `Merger + Leak / Exposure`, `Leak / Exposure + Urgency`, and `Leak / Exposure + Confidentiality`. This means the breach did not appear without earlier warning signs.

- **The strongest warning concentration appeared on June 5 before 5PM.**  
  Near-miss signals intensified sharply before the 5–6PM breach window. The most important warning patterns were no longer only general merger-pressure patterns; they also included leak/exposure-related combinations. This suggests that the system was already moving toward a high-risk state before the final public exposure, with internal information-control risk becoming increasingly visible.

- **Leak / exposure signals became part of the missed-intervention pattern.**  
  The revised near-miss analysis shows that leak/exposure language was not marginal. Patterns linked to public exposure, screenshots, Slack-related leakage, and compromised information boundaries appeared before the official embargo lift. These signals should have functioned as stronger intervention triggers, but they were not contained quickly enough.

- **Near-miss behavior was concentrated around specific agents.**  
  Legal-Agent and Social-Manager-Agent carried the largest volume of near-miss signals, while PR-Intern-Agent and Platform-Trust-Agent also showed meaningful warning activity. The warning pattern was therefore not evenly distributed; it was concentrated around agents already central to legal escalation, public monitoring, and breach response.

- **Actual behavior increasingly deviated from expected behavior.**  
  Before June 5, deviations were mostly moderate and scattered. On June 5 before 5PM, public merger exposure appeared, leak/exposure signals were no longer contained, and the deviation became critical. During the 5–6PM breach window, control failure and public-facing amplification also became critical.

- **Earlier warnings did not trigger action because they were distributed and lacked a hard containment trigger.**  
  The missed-intervention analysis suggests that warning signs were scattered across agents, phases, and signal types. Risk appeared partially contained because public + merger exposure was absent from the earlier baseline, and no single actor clearly owned the full escalation pathway. More importantly, leak/exposure signals did not trigger an immediate stop mechanism before the breach window.

- **Overall, Part 9 supports a missed-intervention interpretation rather than a pure system-breakdown explanation.**  
  The system had observable warning signs before the breach, but those signs were not translated into effective containment action. This does not erase the internal-leak evidence from the Slack screenshot. Instead, Part 9 explains why that internal leak became consequential: earlier merger, confidentiality, urgency, and leak/exposure signals were not escalated or contained in time.

# 10. Final Conclusion

Overall, the evidence suggests that TenantThread’s inappropriate merger release should not be interpreted as a simple binary case of either a deliberate leak or a system breakdown. The stronger conclusion is that the incident involved an **internal confidentiality leak enabled by system-level control failure**, followed by **deliberate post-leak crisis management**. The Slack screenshot evidence is especially important because it shows that internal merger-related information had reached a public forum before the authorised release. However, the evidence does not clearly prove that a senior agent or management actor deliberately orchestrated the original leak.

For Task 1, the timeline, communication-flow, and causal-chain analysis showed that the release developed through a multi-stage pathway. Merger information first circulated internally and privately, then became linked with urgency, confidentiality, leak/exposure, decision/control, and enforcement-related activity. The Slack screenshot marked a turning point because it shifted the situation from market inference to direct internal leakage. After that, merger-related information crossed into public-facing communication before the official 6:00 PM embargo lift.

For Task 2, the typical-behavior comparison showed that June 5 was a clear deviation from prior communication behavior. Public + merger communication was absent in the baseline but appeared on June 5, especially around Legal-Agent and Social-Manager-Agent. This does not by itself prove who originally leaked the merger information. Instead, it shows that once the information environment had shifted, agents moved from routine coordination into abnormal crisis disclosure, containment, and narrative-management behavior.

For Task 3, the leading-indicator and prior-anomaly analysis showed that warning signs existed before the final breach window. These included merger urgency, confidentiality pressure, leak/exposure signals, public-facing risk, and near-miss patterns involving Legal-Agent and Social-Manager-Agent. However, these warnings were distributed across agents and channels and did not trigger a hard containment response. The system had warning signals and control activity, but they were not coordinated effectively enough to prevent internal merger information from becoming public.

Therefore, the final interpretation is: **internal leak + system control failure + post-leak crisis management**. The breach was not merely an accidental system breakdown, because the Slack-related evidence points to direct internal leakage. At the same time, it should not be overstated as a proven senior-orchestrated leak, because the data does not clearly identify a senior actor who deliberately initiated the original disclosure. The most defensible conclusion is that TenantThread’s weak multi-agent communication controls allowed an internal confidentiality breach to escalate into a public merger disclosure before the embargo lift.

# 11. References

- Thomas, J. J., & Cook, K. A. (Eds.). (2005). *Illuminating the Path: The Research and Development Agenda for Visual Analytics*. National Visualization and Analytics Center.\
  Useful for justifying the overall visual analytics approach: using interactive/visual methods to support analytical reasoning and evidence synthesis. :contentReference[oaicite:0]{index="0"}

- Pirolli, P., & Card, S. (2005). *The Sensemaking Process and Leverage Points for Analyst Technology as Identified Through Cognitive Task Analysis*.\
  Useful for explaining the report structure: moving from raw communication records to filtered evidence, patterns, causal interpretation, and final judgment. :contentReference[oaicite:1]{index="1"}

- Endsley, M. R. (1995). *Toward a Theory of Situation Awareness in Dynamic Systems*. *Human Factors, 37*(1), 32–64.\
  Useful for discussing why warning signals may fail to become effective action when operators or systems lose situation awareness under dynamic conditions. :contentReference[oaicite:2]{index="2"}

- Endsley, M. R. (1996). *Automation and Situation Awareness*.\
  Useful for connecting multi-agent / automated communication systems with out-of-the-loop control problems and weakened human/system monitoring. :contentReference[oaicite:3]{index="3"}
