---
title: "Hands-on_Ex06"
author: "HU YUE"
date: "2026-06-05"
date-modified: "last-modified"
format:
  html:
    theme: cosmo
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: true
    code-tools: true
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

# 27 Chapter 27: Modelling, Visualising and Analysing Network Data with R

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts, 
               concaveman, ggforce)
```

This code loads the required R packages for data wrangling, time processing, network analysis, and network visualisation. These packages will be used later to prepare the communication data, construct network objects, and create both static and interactive network graphs.

## 27.3.3 Importing network data from files

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

This code imports the node and edge datasets from the `data` folder. The node file contains information about the people or entities in the email network, while the edge file contains the connections or email interactions between them.

## 27.3.4 Reviewing the imported data

```{r}
glimpse(GAStech_edges)
```

## 27.3.5 Wrangling time

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

## 27.3.7 Wrangling attributes

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

## 27.4 Creating network objects using tidygraph

## 27.4.3 Using tbl_graph() to build tidygraph data model

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

This code creates a directed network graph using the node data and the aggregated edge data. The graph represents email relationships, where the direction of each edge shows who sent emails to whom.

## 27.4.4 Reviewing the output tidygraph’s graph object

```{r}
GAStech_graph
```

## 27.4.6 Changing the active object

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

This code activates the edge table in the network graph and sorts the email connections by `Weight` in descending order. This helps identify the strongest or most frequent email relationships in the network.

## 27.5.1 Plotting a basic network graph

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

This code creates a basic network graph from `GAStech_graph`. The lines represent email connections, and the points represent the nodes in the network.

## 27.5.2 Changing the default network graph theme

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

This code creates a network graph object and applies `theme_graph()` to remove unnecessary chart elements. The result is a cleaner visualisation of the email network.

## 27.5.3 Changing the coloring of the plot

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = 'grey50')) +
  geom_node_point(aes(colour = 'grey40'))

g + theme_graph(background = 'grey10',
                text_colour = 'white')
```

This code creates a darker network visualisation by setting grey colours for the edges and nodes, and applying a dark background with white text. This improves contrast and makes the email network easier to view.

## 27.5.5 Fruchterman and Reingold layout

```{r}
g <- ggraph(GAStech_graph, 
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

This code uses the Fruchterman-Reingold layout to create a network graph. This layout places closely connected nodes nearer to each other, making the structure of the email network easier to interpret.

## 27.5.6 Modifying network nodes

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      size = 3))

g + theme_graph()
```

This code visualises the email network using the `nicely` layout. The nodes are coloured by department, which helps compare communication patterns across different organisational groups.

## 27.5.7 Modifying edges

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width = Weight), 
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

This code visualises the email network with edge width mapped to `Weight`. Thicker lines represent stronger or more frequent email connections, while node colours show different departments.

## 27.6.1 Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width = Weight), 
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)
```

This code applies a graph style and visualises the email network by weekday. The edge width represents the weight of communication, while the faceting separates the network into different weekday panels.

## 27.6.2 Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2) +
  theme(legend.position = 'bottom')
  
g + facet_edges(~Weekday)
```

## 27.6.3 A framed facet graph

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width = Weight), 
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = "bottom")
```

This code creates a faceted email network by weekday. Edge width shows communication weight, node colour shows department, and the foreground theme improves panel visibility while placing the legend at the bottom.

## 27.6.4 Working with facet_nodes()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

## 27.7.1 Computing centrality indices

```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width = Weight), 
                 alpha = 0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
                      size = betweenness_centrality))

g + theme_graph()
```

This code calculates betweenness centrality for each node and uses it to control node size in the network graph. Larger nodes represent people who are more important as bridges between different parts of the email network.

## 27.7.2 Visualising network metrics

```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department, 
                      size = centrality_betweenness()))
g + theme_graph()
```

## 27.7.3 Visualising Community

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(
    group_edge_betweenness(
      weights = Weight, 
      directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(
    aes(
      width=Weight), 
    alpha=0.2) +
  scale_edge_width(
    range = c(0.1, 5)) +
  geom_node_point(
    aes(colour = community))  

g + theme_graph()
```

```{r}
g <- GAStech_graph %>%
  activate(nodes) %>%
  mutate(community = as.factor(
    group_optimal(weights = Weight)),
         betweenness_measure = centrality_betweenness()) %>%
  ggraph(layout = "fr") +
  geom_mark_hull(
    aes(x, y, 
        group = community, 
        fill = community),  
    alpha = 0.2,  
    expand = unit(0.3, "cm"),  # Expand
    radius = unit(0.3, "cm")  # Smoothness
  ) + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(fill = Department,
                      size = betweenness_measure),
                      color = "black",
                      shape = 21)
  
g + theme_graph()
```

## 27.8.1 Data preparation

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```

## 27.8.2 Plotting the first interactive network graph

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated)
```

## 27.8.3 Working with layout

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr")
```

This code creates an interactive network graph using `visNetwork`. The Fruchterman-Reingold layout is applied to position closely connected nodes near each other.

## 27.8.4 Working with visual attributes - Nodes

```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group = Department) 
```

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

This code creates an interactive network graph with a Fruchterman-Reingold layout. It also adds a legend and uses a fixed random seed so that the network layout remains consistent each time the code is rendered.

## 27.8.5 Working with visual attributes - Edges

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to", 
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

## 27.8.6 Interactivity

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

# 29 Visualising and Analysing Text Data with R: tidytext methods

## 29.2.1 Installing and launching R packages

```{r}
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms,
tidyverse, tidygraph, ggraph, igraph)
```

## 29.3 Importing Multiple Text Files from Multiple Folders

```{r}
news20 <- "data/20news/"
```

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}
```

## 29.4 Importing Multiple Text Files from Multiple Folders

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}

raw_text <- tibble(folder = dir(news20, 
                                full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)

write_rds(raw_text, "data/news20.rds")

glimpse(raw_text)
names(raw_text)
```

```{r}
theme_set(theme_gray())
```
## 29.5 Initial EDA

```{r}
raw_text <- read_rds("data/news20.rds")

raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(messages, newsgroup)) +
  geom_col(fill = "lightblue") +
  labs(y = NULL)
```

This code groups the dataset by newsgroup, counts the number of unique messages in each group, and visualises the message counts using a horizontal bar chart.

## 29.6 Introducing tidytext

## 29.6.1 Removing header and automated email signitures

```{r}
cleaned_text <- raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(
           text, "^--")) == 0) %>%
  ungroup()
```

## 29.6.2 Removing lines with nested text representing quotes from other users.

```{r}
cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]")
         | text == "",
         !str_detect(text, 
                     "writes(:|\\.\\.\\.)$"),
         !str_detect(text, 
                     "^In article <")
  )
```

## 29.6.3 Text Data Processing

```{r}
usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
```

This code tokenises the cleaned text into individual words and removes common stop words. It keeps only valid word-like tokens, preparing the text data for further word frequency or text analysis.

```{r}
usenet_words %>%
  count(word, sort = TRUE)
```

```{r}
words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()
```

## 29.6.4 Visualising Words in newsgroups

```{r, fig.width=10, fig.height=8}
wordcloud(
  words_by_newsgroup$word,
  words_by_newsgroup$n,
  max.words = 150,
  scale = c(3, 0.5),
  random.order = FALSE
)
```

This word cloud shows the most frequent words in the text data. The maximum number of words is reduced and the scaling is adjusted so that large words can fit within the plotting area.

## 29.6.5 Visualising Words in newsgroups

```{r, fig.width=10, fig.height=6}
set.seed(1234)

words_by_newsgroup %>%
  filter(newsgroup %in% unique(newsgroup)[1:4]) %>%
  group_by(newsgroup) %>%
  slice_max(n, n = 30, with_ties = FALSE) %>%
  ungroup() %>%
  ggplot(aes(label = word,
             size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  facet_wrap(~newsgroup)
```

This code creates faceted word clouds by newsgroup. The word size represents frequency, and `set.seed(1234)` makes the word cloud layout reproducible each time the code is rendered.

## 29.7 Basic Concept of TF-IDF

## 29.7.1 Computing tf-idf within newsgroups

```{r}
words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE)

tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))
```

## 29.7.2 Visualising tf-idf as interactive table

```{r}
library(DT)

tf_idf %>%
  datatable(
    filter = "top",
    options = list(
      pageLength = 10,
      autoWidth = TRUE
    )
  )
```

This code displays the `tf_idf` dataset as an interactive table. The table allows users to search, sort columns, filter values, and choose how many rows to display.

## 29.7.3 Visualising tf-idf as interactive table

```{r}
DT::datatable(tf_idf, filter = "top") %>% 
  formatRound(columns = c("tf", "idf", "tf_idf"), 
              digits = 3) %>%
  formatStyle(0, 
              target = "row", 
              lineHeight = "25%")
```

This code displays the `tf_idf` table as an interactive table with column filters. The `tf`, `idf`, and `tf_idf` columns are rounded to three decimal places, and the row height is reduced to make the table more compact.

## 29.7.4 Visualising tf-idf within newsgroups

```{r}
tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, 
            n = 12) %>%
  ungroup() %>%
  mutate(word = reorder(word, 
                        tf_idf)) %>%
  ggplot(aes(tf_idf, 
             word, 
             fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ newsgroup, 
             scales = "free") +
  labs(x = "tf-idf", 
       y = NULL)
```

This code filters the TF-IDF results to include only science-related newsgroups. It then selects the top 12 words with the highest TF-IDF score in each group and visualises them as faceted bar charts, showing the most distinctive words for each science newsgroup.

## 29.7.5 Counting and correlating pairs of words with the widyr package

```{r}
newsgroup_cors <- words_by_newsgroup %>%
  pairwise_cor(newsgroup, 
               word, 
               n, 
               sort = TRUE)
```

## 29.7.6 Visualising correlation as a network

```{r}
set.seed(2017)

newsgroup_cors %>%
  filter(correlation > .025) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation, 
                     width = correlation)) +
  geom_node_point(size = 6, 
                  color = "lightblue") +
  geom_node_text(aes(label = name),
                 color = "red",
                 repel = TRUE) +
  theme_void()
```

This code visualises the correlation network between newsgroups. Only correlations above 0.025 are kept, and stronger correlations are shown with thicker and more visible edges. The nodes represent newsgroups, while the labels identify each group in the network.

## 29.7.7 Bigram

```{r}
bigrams <- cleaned_text %>%
  unnest_tokens(bigram, 
                text, 
                token = "ngrams", 
                n = 2)
bigrams
```

## 29.7.8 Counting bigrams

```{r}
bigrams_count <- bigrams %>%
  filter(bigram != 'NA') %>%
  count(bigram, sort = TRUE)

bigrams_count

```

## 29.7.9 Cleaning bigram

```{r}
bigrams_separated <- bigrams %>%
  filter(bigram != "NA") %>%
  separate(bigram, c("word1", "word2"), 
           sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
bigrams_filtered
```

This code separates each bigram into two individual words and removes bigrams that contain stop words. The filtered result keeps more meaningful word pairs for further bigram analysis.

## 29.7.10 Counting the bigram again

```{r}
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)
```

## 29.7.11 Create a network graph from bigram data frame

```{r}
bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph
```

## 29.7.12 Visualizing a network of bigrams with ggraph

```{r}
set.seed(1234)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1)
```

This code visualises the bigram network using a Fruchterman-Reingold layout. The edges represent connections between words that appear together as bigrams, while the nodes represent individual words.

## 29.7.13 Revised version

```{r}
set.seed(1234)

a <- grid::arrow(type = "closed", 
                 length = unit(.15,
                               "inches"))

ggraph(bigram_graph, 
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), 
                 show.legend = FALSE,
                 arrow = a, 
                 end_cap = circle(.07,
                                  'inches')) +
  geom_node_point(color = "lightblue", 
                  size = 5) +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1) +
  theme_void()
```
