---
title: "Hands-on_Ex02"
aurthor: HU Yue
format: html
---

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```
## 2.2.2 Importing data
```{r}
library(readr)
exam_data <- readr::read_csv("Exam_data.csv")

```
## 2.3 Beyond ggplot2 Annotation: ggrepel

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

**Explanation**

This code creates a scatter plot of `MATHS` versus `ENGLISH`, adds a linear regression line, and labels each point using `ID`. `geom_label_repel()` helps prevent overlapping labels, while `coord_cartesian()` restricts the axes to the score range of 0 to 100.

## 2.4 Beyond ggplot2 Themes
```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

**Explanation**

This code creates a histogram of `MATHS` scores. The data is grouped into 20 bins, and `boundary = 100` aligns the bins at a fixed boundary. The bars are styled with grey borders and light grey fill, and `theme_gray()` applies the default ggplot theme for a clean appearance.

## 2.4.1 Working with ggtheme package

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

**Explanation**

This code creates a histogram of `MATHS` scores using 20 bins and aligns them with a fixed boundary. The bars are styled with grey borders and light grey fill. `theme_economist()` applies a professional theme inspired by *The Economist*, giving the chart a more polished and publication-style appearance.

## 2.4.2 Working with hrbthems package

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

**Explanation**

This code creates a histogram of `MATHS` scores using 20 bins and aligns them with a fixed boundary. The bars are styled with grey borders and light grey fill. `theme_ipsum()` applies a clean, modern theme with improved typography, making the chart more visually appealing.

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

**Explanation**

This code creates a histogram of `MATHS` scores and applies a customized `theme_ipsum()`. The theme increases the axis title size and overall text size for better readability, while `grid = "Y"` keeps only horizontal grid lines. This results in a cleaner and more polished visualization.

## 2.5 Beyond Single Graph


```{r}
p1 <- ggplot(data = exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") + 
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of Maths scores")

p1
```

**Explanation**

This code creates a histogram of `MATHS` scores and assigns it to `p1`. `coord_cartesian(xlim = c(0, 100))` restricts the visible x-axis range to 0–100 without removing any data. Displaying `p1` renders the plot.

```{r}
p2 <- ggplot(data = exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins = 20, 
                 boundary = 100,
                 color = "grey25", 
                 fill = "grey90") +
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of English scores")

p2
```

**Explanation**

This code creates a histogram of `ENGLISH` scores and assigns it to `p2`. The data is grouped into 20 bins, and `coord_cartesian(xlim = c(0, 100))` limits the visible x-axis range to 0–100 without removing data. Displaying `p2` renders the plot.


```{r}
p3 <- ggplot(data = exam_data, 
             aes(x = MATHS, 
                 y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              linewidth = 0.5) +  
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

p3
```

**Explanation**

This code creates a scatter plot of `MATHS` versus `ENGLISH` and adds a linear regression line. `geom_point()` shows individual observations, while `geom_smooth(method = lm)` fits a straight line to indicate the relationship between the two variables. `coord_cartesian()` limits both axes to the 0–100 score range for clearer interpretation.

## 2.5.1 Creating Composite Graphics: pathwork methods

## 2.5.2 Combining two ggplot2 graphs

```{r}
p1 + p2
```

**Explanation**

This code attempts to combine `p1` and `p2`. However, in `ggplot2`, plots cannot be directly added together using `+`. To properly combine multiple plots, packages such as `patchwork` or `gridExtra` should be used.

## 2.5.3 Combining three ggplot2 graphs

```{r}
library(patchwork)

(p1 / p2) | p3
```

**Explanation**

This code uses the `patchwork` package to arrange multiple plots. `p1 / p2` stacks `p1` above `p2` vertically, and `| p3` places `p3` to the right. The result is a combined layout with `p1` and `p2` on the left (top and bottom) and `p3` on the right.

## 2.5.4 Creating a composite figure with tag

```{r}
library(patchwork)

((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

**Explanation**

This code arranges the plots using `patchwork` and adds automatic labels to each subplot. `p1 / p2` stacks `p1` above `p2`, and `| p3` places `p3` to the right. `plot_annotation(tag_levels = 'I')` labels the plots as I, II, and III, making it easier to reference each subplot.

## 2.5.5 Creating figure with insert


```{r}
library(patchwork)

p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

**Explanation**

This code overlays `p2` onto `p3` as an inset plot using the `patchwork` package. The arguments `left`, `bottom`, `right`, and `top` control the position and size of the inset using relative coordinates (from 0 to 1). This allows a secondary plot to be displayed within the main plot for comparison or emphasis.

## 2.5.6 Creating a composite figure by using patchwork and ggtheme

```{r}
library(patchwork)
library(ggthemes)

patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

**Explanation**

This code first combines `p1`, `p2`, and `p3` into a layout using `patchwork`. The operator `&` is then used to apply `theme_economist()` to all plots in the layout at once. This ensures a consistent visual style across the combined plots.