---
title: "Hands-on_Ex05"
author: "HU YUE"
date: "2026-05-13"
date-modified: "last-modified"
format: html

execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## 13  Creating Ternary Plot with R

## 13.2 Installing and launching R packages

```{r}
pacman::p_load(plotly, ggtern, tidyverse)
```

## 13.3 Data Preparation

```{r}
#Reading the data into R environment
pop_data <- read_csv("respopagsex2000to2018_tidy.csv") 
```

## 13.3.3 Preparing the Data

```{r}
agpop_mutated <- pop_data %>%
  mutate(Year = as.character(Year)) %>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8])) %>%
  mutate(ACTIVE = rowSums(.[9:16])) %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018) %>%
  filter(TOTAL > 0)
```

**Explanation:**  
This code reshapes the population data from long format to wide format by age group.  
It then calculates the young, active, old, and total population groups for 2018, keeping only areas with valid population values.

## 13.4 Plotting Ternary Diagram with R
## 13.4.1 4.1 Plotting a static ternary diagram

```{r}
ggtern(data = agpop_mutated, aes(x = YOUNG, y = ACTIVE, z = OLD)) +
  geom_point()
```

**Explanation:**  
This code creates a ternary plot showing the proportions of young, active, and old populations. Each point represents one area’s population structure across the three age groups.

```{r}
# Reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt,
    tickformat = ".0%",
    tickfont = list(size = 10),
    titlefont = list(size = 13)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young Population"),
  baxis = axis("Active Population"),
  caxis = axis("Old Population")
)

plot_ly(
  agpop_mutated,
  a = ~YOUNG / TOTAL,
  b = ~ACTIVE / TOTAL,
  c = ~OLD / TOTAL,
  type = "scatterternary",
  mode = "markers",
  marker = list(
    size = 8,
    color = "#2E86DE",
    opacity = 0.65,
    line = list(color = "white", width = 1)
  ),
  text = ~paste(
    "Young:", round(YOUNG / TOTAL * 100, 1), "%",
    "<br>Active:", round(ACTIVE / TOTAL * 100, 1), "%",
    "<br>Old:", round(OLD / TOTAL * 100, 1), "%"
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = list(
      text = "Interactive Population Structure by Age Group",
      y = 0.95
    ),
    margin = list(t = 90, l = 40, r = 40, b = 40),
    annotations = list(
      list(
        text = "Population Composition",
        x = 0.1, y = 0.93,
        ax = 0, ay = 0,
        xref = "paper", yref = "paper",
        align = "center",
        font = list(family = "serif", size = 14, color = "white"),
        bgcolor = "#2C3E50",
        bordercolor = "black",
        borderwidth = 2
      )
    ),
    ternary = ternaryAxes
  )
```

**Explanation:**  
The error happened because `ternaryAxes` was used before it was created.  
This version defines the axis formatting first, so the ternary plot can render correctly.

## 6  Visual Correlation Analysis

## 6.2 Installing and Launching R Packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

## 6.3 Importing and Preparing The Data Set
## 6.3.1 Importing Data

```{r}
wine <- read_csv("wine_quality.csv")
```

## 6.4 Building Correlation Matrix: pairs() method
## 6.4.1 Building a basic correlation matrix

```{r}
pairs(wine[, 1:11])
```

**Explanation:**  
This code generates a scatterplot matrix for the first 11 variables in the wine dataset.  
It helps visualize pairwise relationships and detect correlations or patterns across variables.

```{r}
pairs(wine[,2:12])
```

## 6.4.2 Drawing the lower corner

```{r}
pairs(wine[, 2:12], upper.panel = NULL)
```

**Explanation:**  
This code creates a lower-triangle scatterplot matrix for variables 2 to 12 in the wine dataset.  
Removing the upper panel makes the plot cleaner and easier to read.

```{r}
pairs(wine[, 2:12], lower.panel = NULL)
```

**Explanation:**  
This code displays only the upper-triangle scatterplot matrix for variables 2 to 12.  
Hiding the lower panel reduces redundancy and improves visual clarity.

## 6.4.3 Including with correlation coefficients- change color, hover, adding 
```{r}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  
  r <- cor(x, y, use = "complete.obs")
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  
  if (missing(cex.cor)) {
    cex.cor <- 0.8 / strwidth(txt)
  }
  
  text(
    0.5, 0.5, txt,
    cex = cex.cor * (1 + abs(r)) / 2,
    col = "#1F4E79",
    font = 2
  )
}

panel.smooth2 <- function(x, y, ...) {
  points(x, y, pch = 19, col = rgb(0.2, 0.4, 0.7, 0.35), cex = 0.6)
  lines(lowess(x, y), col = "#D95F02", lwd = 2)
}

pairs(
  wine[, 2:12],
  lower.panel = panel.smooth2,
  upper.panel = panel.cor,
  diag.panel = NULL,
  main = "Scatterplot Matrix and Correlation of Wine Variables",
  pch = 19
)
```

**Explanation:**  
This improved scatterplot matrix shows smoothed trends in the lower panel and correlation values in the upper panel.  
The transparent points and colored trend lines make relationships between wine variables easier to interpret.

## 6.5 Visualising Correlation Matrix: ggcormat() - changing color

## 6.5.1 The basic plot

```{r}
ggstatsplot::ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  colors = c("red", "white", "green")
)
```

**Explanation:**  
This code customizes the correlation matrix colors to a red–white–green gradient.  
Red indicates negative correlation, green indicates positive correlation, and white represents near zero correlation.

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  colors = c("red", "white", "green"),
  ggcorrplot.args = list(
    outline.color = "black", 
    hc.order = TRUE,
    tl.cex = 3
  ),
  title = "Correlogram for wine dataset",
  subtitle = "Four pairs are not significant at p < 0.05"
)
```

**Explanation:**  
This code updates the correlogram to use a red–white–green color scheme.  
It keeps the clustering and styling options while making positive and negative correlations visually distinct.

## 6.6 Building multiple plots

```{r fig.width=14, fig.height=10}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  cor.vars.names = c(
    "FA", "VA", "CA", "RS", "Cl", "FSO2",
    "TSO2", "Den", "pH", "Sul", "Alc"
  ),
  grouping.var = type,
  matrix.type = "lower",
  type = "robust",
  p.adjust.method = "holm",
  digits = 1,
  colors = c("red", "white", "green"),
  plotgrid.args = list(ncol = 1),
  ggcorrplot.args = list(
    outline.color = "black",
    hc.order = TRUE,
    pch.cex = 10
  ),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "Grouped by wine type",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```
**Explanation:**  
This code creates grouped correlation matrices by wine type using robust correlation.  
The red–white–green color scale makes negative, weak, and positive correlations easier to compare across groups.


## 6.7 Visualising Correlation Matrix using corrplot Package

## 6.7.1 Getting started with corrplot

```{r}
wine.cor <- cor(wine[, 1:11])

corrplot(wine.cor)
```

## 6.7.2 Working with visual geometrics

```{r}
corrplot(wine.cor, 
         method = "ellipse")
```

**Explanation:**  
This code creates a correlation plot using ellipses to represent the strength and direction of correlations.  
The shape and orientation of each ellipse help visualize whether the relationship is positive, negative, weak, or strong.

## 6.7.3 Working with layout

```{r}
corrplot(
  wine.cor,
  method = "ellipse",
  type = "lower"
)
```

**Explanation:**  
This code creates a correlation plot with ellipses and displays only the lower triangle of the matrix.  
Showing only half of the matrix reduces repetition and makes the plot easier to read.

```{r}
corrplot(
  wine.cor,
  method = "ellipse",
  type = "lower",
  diag = FALSE,
  tl.col = "black"
)
```

**Explanation:**  
This code creates a lower-triangle correlation plot using ellipses and removes the diagonal elements.  
The variable labels are shown in black to improve readability.

## 6.7.4 Working with mixed layout

```{r}
corrplot.mixed(
  wine.cor,
  lower = "ellipse",
  upper = "number",
  tl.pos = "lt",
  diag = "l",
  tl.col = "black"
)
```

**Explanation:**  
This code creates a mixed correlation plot with ellipses in the lower triangle and correlation numbers in the upper triangle.  
This combined format makes it easy to compare both the visual pattern and the exact correlation values.

```{r}
corrplot.mixed(
  wine.cor,
  lower = "ellipse",
  upper = "number",
  tl.pos = "lt",
  diag = "l",
  tl.col = "black"
)
```

**Explanation:**  
This code produces a mixed correlation plot with ellipses in the lower triangle and numerical correlation values in the upper triangle.  
It combines visual interpretation and exact values, making the correlation structure easier to understand.

## 6.7.5 Combining corrgram with the significant test

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

## 6.7.6 Reorder a corrgram

```{r}
corrplot.mixed(
  wine.cor,
  lower = "ellipse",
  upper = "number",
  tl.pos = "lt",
  diag = "l",
  order = "AOE",
  tl.col = "black"
)
```

**Explanation:**  
This code creates a mixed correlation plot with ellipses in the lower triangle and correlation values in the upper triangle.  
The `order = "AOE"` option rearranges variables based on angular order of eigenvectors, which helps reveal clusters and correlation patterns more clearly.

## 6.7.7 Reordering a correlation matrix using hclust

```{r}
corrplot(
  wine.cor,
  method = "ellipse",
  tl.pos = "lt",
  tl.col = "black",
  order = "hclust",
  hclust.method = "ward.D",
  addrect = 3
)
```

**Explanation:**  
This code creates an ellipse-based correlation plot and reorders the variables using hierarchical clustering.  
The rectangles highlight three variable clusters, making the correlation structure easier to interpret.

## 14  Heatmap for Visualising and Analysing Multivariate Data

## 14.2 Installing and Launching R Packages

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## 14.3 Importing and Preparing The Data Set
```{r}
wh <- read_csv("WHData-2018.csv")
```

### 14.3.2 Preparing the data
```{r}
row.names(wh) <- wh$Country
```

## 14.3.3 Transforming the data frame into a matrix
```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## 14.4 Static Heatmap
## 14.4.1 heatmap() of R Stats - changing color more readable

```{r}
wh_heatmap <- heatmap(
  wh_matrix,
  Rowv = NA,
  Colv = NA
)
```

**Explanation:**  
This code creates a heatmap from the `wh_matrix` data matrix.  
Setting `Rowv = NA` and `Colv = NA` removes row and column clustering, so the heatmap keeps the original ordering of the data.

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

**Explanation:**  
This code creates a heatmap from `wh_matrix` and applies the default hierarchical clustering to both rows and columns.  
It helps reveal patterns, similarities, and possible clusters in the data.

```{r}
wh_heatmap <- heatmap(
  wh_matrix,
  scale = "column",
  col = colorRampPalette(c("navy", "white", "firebrick3"))(50),
  cexRow = 0.7,
  cexCol = 0.9,
  margins = c(10, 5),
  Colv = NA
)
```

**Explanation:**  
This code improves the heatmap by using a stronger blue–white–red color contrast.  
It also removes column clustering to preserve the original column order and makes the labels easier to read.

## 14.5 Creating Interactive Heatmap

## 14.5.1 Working with heatmaply

```{r}
heatmaply(
  mtcars,
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100)
)
```

**Explanation:**  
This code creates an interactive heatmap of the `mtcars` dataset with a stronger blue–white–red color palette.  
The customized colors make low, medium, and high values easier to distinguish.

```{r}
heatmaply(
  wh_matrix[, -c(1, 2, 4, 5)],
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 6,
  fontsize_col = 6
)
```

**Explanation:**  
This code creates an interactive heatmap with a clearer blue–white–red color scale and smaller row and column labels.  
Reducing the font size makes the variable names easier to fit and prevents the text from overlapping.

## 14.5.2 Data trasformation

```{r}
heatmaply(
  wh_matrix[, -c(1, 2, 4, 5)],
  scale = "column",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Interactive Heatmap of Selected Variables"
)
```

**Explanation:**  
This code creates a column-scaled interactive heatmap for the selected variables in `wh_matrix`.  
The improved color palette and smaller text make the heatmap easier to read and interpret.

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 6,
  fontsize_col = 6,
  main = "Interactive Heatmap of Normalized Variables"
)
```

**Explanation:**  
This code creates an interactive heatmap after normalizing the selected variables, so they are on a comparable scale.  
The stronger color palette and smaller labels make patterns across variables easier to see.

```{r}
heatmaply(
  percentize(wh_matrix[, -c(1, 2, 4, 5)]),
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Interactive Heatmap of Percentized Variables"
)
```

**Explanation:**  
This code creates an interactive heatmap after converting the selected variables to percentage-based values.  
The customized color palette and smaller text improve readability and make patterns easier to compare across variables.

## 14.5.4 Manual approach

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  dist_method = "euclidean",
  hclust_method = "ward.D",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 6,
  fontsize_col = 6,
  main = "Normalized Heatmap with Hierarchical Clustering"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized values, Euclidean distance, and Ward’s clustering method.  
The customized color scale and smaller labels make the clustered patterns easier to read and compare.

## 14.5.5 Statistical approach

```{r}
wh_d <- dist(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  method = "euclidean"
)

dend_expend(wh_d)[[3]]
```

**Explanation:**  
This code first computes the Euclidean distance matrix based on the normalized variables in `wh_matrix`.  
It then extracts the third component from `dend_expend(wh_d)`, which is typically used to inspect part of the dendrogram expansion results.

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

**Explanation:**  
This code performs hierarchical clustering on the distance matrix using the average linkage method.  
It then uses `find_k()` to help identify a suitable number of clusters and plots the result for visual inspection.

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  dist_method = "euclidean",
  hclust_method = "average",
  k_row = 3,
  scale = "none",
  colors = colorRampPalette(c("darkblue", "white", "darkred"))(120),
  fontsize_row = 7,
  fontsize_col = 8,
  margins = c(80, 120),
  main = "Clustered Heatmap of Normalized Variables"
)
```

**Explanation:**  
This version enhances the clustered heatmap with stronger color contrast and better spacing.  
It makes the row clusters easier to distinguish while keeping the normalized values comparable across variables.

## 14.5.6 Seriation

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  seriate = "OLO",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Normalized Heatmap with Optimal Leaf Ordering"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized data and applies optimal leaf ordering (`OLO`) to improve the arrangement of rows and columns.  
The customized color palette and smaller labels make the reordered patterns easier to see.

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  seriate = "GW",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Normalized Heatmap with Gruvaeus-Wainer Reordering"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized data and applies the Gruvaeus–Wainer (`GW`) seriating method to reorder the rows and columns.  
The improved color contrast and smaller labels make the reordered structure easier to read.

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  seriate = "mean",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Normalized Heatmap with Mean Reordering"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized data and reorders the rows and columns based on the mean values.  
The stronger color palette and smaller labels make the reordered patterns easier to compare and interpret.

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  seriate = "none",
  colors = colorRampPalette(c("navy", "white", "firebrick3"))(100),
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Normalized Heatmap without Reordering"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized data without reordering the rows or columns.  
It keeps the original data order while improving readability with a clearer color scale and smaller labels.

## 14.5.7 Working with colour palettes

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  seriate = "none",
  colors = Blues,
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Normalized Heatmap with Blue Colour Palette"
)
```

**Explanation:**  
This code creates an interactive heatmap using normalized data with no reordering of rows or columns.  
The `Blues` palette gives the heatmap a cleaner single-colour gradient, while the smaller text improves readability.

## 14.5.8 The finishing touch

```{r}
heatmaply(
  normalize(wh_matrix[, -c(1, 2, 4, 5)]),
  Colv = NA,
  seriate = "none",
  colors = Blues,
  k_row = 5,
  margins = c(60, 220, 80, 40),
  fontsize_row = 3,
  fontsize_col = 7,
  main = "World Happiness Indicators by Country, 2018",
  xlab = "World Happiness Indicators",
  ylab = "World Countries",
  width = 1100,
  height = 1400
)
```

**Explanation:**  
This code creates an interactive heatmap of normalized World Happiness indicators by country.  
The figure size is enlarged, the title is shortened, and the row labels are reduced to improve readability.


## 15 Visual Multivariate Analysis with Parallel Coordinates Plot

## 15.2 Installing and Launching R Packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## 15.3 Data Preparation
```{r}
wh <- read_csv("WHData-2018.csv")
```

## 15.4 Plotting Static Parallel Coordinates Plot
## 15.4.1 Plotting a simple parallel coordinates
```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

## 15.4.2 Plotting a parallel coordinates with boxplot

```{r}
ggparcoord(
  data = wh,
  columns = c(7:12),
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Parallel Coordinates Plot of World Happiness Variables"
) +
  theme(
    axis.text.x = element_text(
      angle = 35,
      hjust = 1,
      vjust = 1,
      size = 10
    ),
    plot.title = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```

**Explanation:**  
This code improves the parallel coordinates plot by rotating the x-axis labels and reducing text crowding.  
It also adjusts the title and legend text sizes to make the overall plot easier to read.

## 15.4.3 Parallel coordinates with facet

```{r}
ggparcoord(
  data = wh,
  columns = c(7:12),
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Multiple Parallel Coordinates Plots of World Happiness Variables by Region"
) +
  facet_wrap(~Region) +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1,
      size = 8
    ),
    strip.text = element_text(size = 10),
    plot.title = element_text(size = 16)
  )
```

**Explanation:**  
This code creates multiple parallel coordinates plots by region using `facet_wrap()`.  
Rotating the x-axis labels and reducing the text size makes the variable names easier to read across the smaller facet panels.

## 15.4.4 Rotating x-axis text label

```{r}
ggparcoord(
  data = wh,
  columns = c(7:12),
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Multiple Parallel Coordinates Plots of World Happiness Variables by Region"
) +
  facet_wrap(~Region) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8)
  )
```

**Explanation:**  
This code creates multiple parallel coordinates plots by region and rotates the x-axis labels by 30 degrees.  
The label adjustment helps reduce overlap and improves readability in the faceted panels.

## 15.4.5 Adjusting the rotated x-axis text label

```{r fig.width=12, fig.height=8}
ggparcoord(
  data = wh,
  columns = c(7:12),
  groupColumn = 2,
  scale = "uniminmax",
  alphaLines = 0.2,
  boxplot = TRUE,
  title = "Multiple Parallel Coordinates Plots of World Happiness Variables by Region"
) +
  facet_wrap(~Region, ncol = 3) +
  theme(
    plot.title = element_text(size = 14),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 6),
    axis.text.y = element_text(size = 7),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    strip.text = element_text(size = 8),
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 7)
  ) +
  guides(colour = guide_legend(nrow = 2))
```

**Explanation:**  
This version controls the facet layout so each panel has more space.  
Smaller text and a wider panel layout make the plot easier to read and closer to a report-ready format.

**Explanation:**  
This code creates faceted parallel coordinates plots for different world regions.  
The x-axis labels are rotated by 30 degrees and right-aligned to reduce overlap and improve readability.

## 15.5 Plotting Interactive Parallel Coordinates Plot: parallelPlot methods
## 15.5.1 The basic plot

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

## 15.5.2 Rotate axis label

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

## 15.5.3 Changing the colour scheme

```{r}
parallelPlot(
  wh,
  continuousCS = "YlOrRd",
  rotateTitle = TRUE
)
```

**Explanation:**  
This code creates a parallel coordinate plot for the `wh` dataset using the `"YlOrRd"` colour scheme.  
The plot title is rotated to improve readability and make the overall layout clearer.

## 15.5.4 Parallel coordinates plot with histogram
```{r}
histoVisibility <- rep(TRUE, ncol(wh))

parallelPlot(
  wh,
  rotateTitle = TRUE,
  histoVisibility = histoVisibility
)
```

**Explanation:**  
This code creates a parallel coordinate plot and enables histograms for all variables in the dataset.  
The rotated titles improve readability, while the histograms provide additional information about the distribution of each variable.

## 16  Treemap Visualisation with R
## 16.2 Installing and Launching R Packages

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

## 16.3 Data Wrangling

```{r}
realis2018 <- read_csv("realis2018.csv")
```

## 16.3.3 Grouped summaries without the Pipe

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

## 16.3.4 Grouped summaries with the pipe
```{r}
realis2018_summarised <- realis2018 %>%
  group_by(
    `Project Name`, `Planning Region`,
    `Planning Area`, `Property Type`,
    `Type of Sale`
  ) %>%
  summarise(
    `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
    `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
    `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
    `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE)
  )
```

**Explanation:**  
This code groups the 2018 real estate data by project and property-related categories.  
It then calculates the total units sold, total area, median unit price per square metre, and median transacted price for each group.

## 16.4 Designing Treemap with treemap Package

## 16.4.1 Designing a static treemap

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

**Explanation:**  
This code filters the summarized real estate dataset to keep only resale condominium transactions.  
It helps focus the analysis on a specific property type and sale category.

## 16.4.2 Using the basic arguments

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap of resale condominium projects grouped by planning region, planning area, and project name.  
The box size represents the total number of units sold, while the colour shows the median unit price per square metre.

## 16.4.3 Working with vColor and type arguments

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "value",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap of resale condominium projects and colours the boxes based on the actual values of median unit price per square metre.  
The box size shows total units sold, while the colour intensity reflects the price level directly.

## 16.4.5 The “value” type treemap

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "value",
  palette = "RdYlBu",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap for resale condominium projects, where box size represents total units sold and box colour represents the median unit price per square metre.  
The `RdYlBu` palette adds a diverging colour scheme, making differences in price levels easier to distinguish.

## 16.4.6 The “manual” type treemap

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "manual",
  palette = "RdYlBu",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap for resale condominium projects, where box size represents the total number of units sold.  
With `type = "manual"`, the colour palette is applied manually to the values of median unit price per square metre.


```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "manual",
  palette = "Blues",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap of resale condominium projects, where the box size represents the total number of units sold.  
The `Blues` palette is used to manually colour the boxes according to the median unit price per square metre.

## 16.4.8 Working with algorithm argument

```{r}
treemap(
  realis2018_selected,
  index = c("Planning Region", "Planning Area", "Project Name"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "manual",
  palette = "Blues",
  algorithm = "squarified",
  title = "Resale Condominium by Planning Region and Area, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap of resale condominium projects, with box size showing total units sold and colour showing the median unit price per square metre.  
The `squarified` algorithm arranges the rectangles into more balanced shapes, making the treemap easier to read.

## 16.4.9 Using sortID

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 16.5 Designing Treemap using treemapify Package

## 16.5.1 Designing a basic treemap

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Unit Sold`,
    fill = `Median Unit Price ($ psm)`
  ),
  layout = "scol",
  start = "bottomleft"
) +
  geom_treemap() +
  scale_fill_gradient(low = "lightblue", high = "blue")
```

**Explanation:**  
This code creates a treemap using `ggplot2`, where the rectangle size represents the total number of units sold.  
The fill colour uses a light-to-dark blue gradient to show differences in median unit price per square metre.

## 16.5.2 Defining hierarchy

```{r}
ggplot(
  data = realis2018_selected,
  aes(
    area = `Total Unit Sold`,
    fill = `Median Unit Price ($ psm)`,
    subgroup = `Planning Region`
  ),
  start = "topleft"
) +
  geom_treemap()
```

**Explanation:**  
This code creates a treemap where the rectangle size shows the total number of units sold and the fill colour represents the median unit price per square metre.  
The `subgroup` aesthetic groups the treemap by planning region, helping to organize the projects into larger regional sections.
## Group by Planning Area
```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap()
```
## Adding boundary line
```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

## 16.6 Designing Interactive Treemap using d3treeR

16.6.1 Installing d3treeR package

```{r}
install.packages("devtools", repos = "https://cloud.r-project.org")
library(devtools)
install_github("timelyportfolio/d3treeR")
library(d3treeR)
```

## 16.6.2 Designing An Interactive Treemap
```{r}
tm <- treemap(
  realis2018_summarised,
  index = c("Planning Region", "Planning Area"),
  vSize = "Total Unit Sold",
  vColor = "Median Unit Price ($ psm)",
  type = "value",
  title = "Private Residential Property Sold, 2017",
  title.legend = "Median Unit Price (S$ per sq. m)"
)
```

**Explanation:**  
This code creates a treemap of private residential property sales grouped by planning region and planning area.  
The box size represents the total number of units sold, while the colour shows the median unit price per square metre.

```{r}
#| out-width: 100%

d3tree(
  tm,
  rootname = "Private Residential Property Sales in Singapore"
)
```

**Explanation:**  
This version keeps the same interactive tree but gives it more horizontal space in the webpage output.  
It helps the labels and hierarchy display more clearly.